---
title: "TidyX 23 - For Loops"
author: "Patrick Ward"
date: "8/15/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(skimr)
library(janitor)

theme_set(theme_light())

```


### Pythagorean Wins Formula for Baseball

Load data from baseball from 1980 - 2006


```{r}

baseball_wins <- read_csv(
  here("TidyTuesday_Explained/023-Mailbag_and_Expected_Wins/Pythagorean Wins.csv")
  ) %>% 
  clean_names()

baseball_wins %>% skim()

```


Calculate the scoring ratio and win percent for each team

```{r}

baseball_ratio <- baseball_wins %>%
  mutate(
    scoring_ratio = runs_scored / runs_allowed,
    win_pct = wins / (wins + losses)
    )

```


Calculate Pythagorean Wins Percent for each team

$PythWinPct = Scoring Ratio^i / (1 + Scoring Ratio^i)$

The exponent `i` will initially be `3` but we will optimize this with our `for` loop.

```{r}

i <- 3

baseball_ratio_pythag <- baseball_ratio %>%
  mutate(
    win_pct_calc = scoring_ratio^i / (1 + scoring_ratio^i)
    )

```


Calculate the absolute error

```{r}
baseball_error <- baseball_ratio_pythag %>%
  mutate(
    error = win_pct - win_pct_calc
    )


## Visual 1
baseball_error %>%
  ggplot(aes(x = error)) +
  geom_histogram(alpha = 0.6) +
  geom_vline(xintercept = 0,
             size = 1.2,
             color = "red",
             linetype = "dashed") +
  ggtitle("Error between Actual Win% and Pythagorean Win%")

## Visual 2
baseball_error %>%
  ggplot() +
  geom_histogram(aes(x = win_pct),
                 fill = "light blue",
                 alpha = 0.6) +
  geom_histogram(aes(x = win_pct_calc),
                 fill = "light green",
                 alpha = 0.6) +
  annotate("text", 
           x = 0.3,
           y = 70,
           label = "Actual Win%",
           color = "light blue",
           size = 6) +
    annotate("text", 
           x = 0.3,
           y = 65,
           label = "Pythagorean Win%",
           color = "light green",
           size = 6) +
  ggtitle("Actual Win% and Pythagorean Win%")

```

Calculate the mean absolute error

```{r}

MAE <- mean(abs(baseball_error$error))
MAE

```


Our Pythagorean Win% estimation is about 3.4% off. Can we find a more optimal exponent than `3`?


### Using a for loop to optimize the Pythagorean Win Formual Exponent

First create a range of possible exponents to try

```{r}

i_options <- seq(from = 0.5, to = 4, by = 0.1)

```


Create an empty data frame for depositing the results of the mean absolute error for each exponent in the sequence

```{r}

mae_results <- data.frame("Exp" = i_options, "Results" = NA)

```

Use a `for` loop to test all of the exponents in the equation

# wholly in base R

```{r}
d2 <- baseball_ratio[, 3:6]
head(d2)
```


Use a `for` loop to test all of the exponents in the equation

```{r}

for(i in 1:length(i_options)){

	win.pct <- d2$wins / (d2$wins + d2$losses)	
	
	pred.win.pct <- (d2$runs_scored/d2$runs_allowed)^i_options[i] / (1 + (d2$runs_scored/d2$runs_allowed)^i_options[i])
	
	mae_results[i,2] <- mean(abs(win.pct - pred.win.pct))
	
}
```

# function and Mixed tidyverse/Base R 

```{r}

## abstract the calculation

pythagorean_wins <- function(scored, allowed, i){
  
  p_win_pct <- (scored/allowed) ^ i / (1 + (scored/allowed) ^ i )
  return(p_win_pct)
  
}


```

```{r}


for(i in 1:length(i_options)){

	baseball_ratio_options <- baseball_ratio %>% 
	  mutate(
	    pred_win_pct = pythagorean_wins(
	      scored = runs_scored,
	      allowed = runs_allowed,
	      i = i_options[[i]]
	      )
	  ) %>% 
	  select(
	    win_pct, pred_win_pct
	  )
	
	mae_results[i,2] <- mean(abs(baseball_ratio_options[[1]]- baseball_ratio_options[[2]]))
	
}
```


Look at the MAE results for each exponent

```{r}

mae_results %>% 
  head()

mae_results %>%
  ggplot(aes(x = Exp, y = Results)) +
  geom_line() +
  labs(x = "Exponent",
       y = "Mean Absolute Error",
       title = "MAE for Different Exponents in the Pythagorean Wins Formula")

```


Obtain the exponent that will minimize the different between the actual and predicted win%

```{r}
min(mae_results$Results) # Lowest MAE is 0.0196

## Get the exponent
mae_results %>% 
  filter(Results == min(Results))

```


## Tidyverse Approach

```{r}

mae_tibble <- seq(from = 0.5, to = 4, by = 0.1) %>%
  
  map_dfr(function(i) {
    
    baseball_ratio_options <- baseball_ratio %>%
      mutate(pred_win_pct = pythagorean_wins(
        scored = runs_scored,
        allowed = runs_allowed,
        i = i
      )) %>%
      select(win_pct, pred_win_pct)
    
    tibble(
      Exp = i,
      Results = mean(abs(baseball_ratio_options[[1]] - baseball_ratio_options[[2]]))
    )
  })

mae_tibble %>% 
  filter(Results == min(Results))


```


