---
title: "Untitled"
author: "Patrick Ward"
date: "10/1/2020"
output: html_document
---

#### Load packages & data

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

### Load Packages -------------------------------------------------
library(tidyverse)
library(tidytext)
library(rtweet)
library(here)

theme_set(theme_light())

```


```{r scrape}

## create path for rds file
debate_tweets_rds <- here("TidyTuesday_Explained/030-Sentiment/debate_tweets_BIG.rds")

## If RDS does not exist, scrape tweets using {rtweet}

if(!file.exists(debate_tweets_rds)){
  
  ### Scrape Twitter for #Debate20202 -----------------------------
  # Get 10,000 tweets, exclude retweets, and only tweets in English
  
  debate_tweets <- search_tweets("#Debate2020",
                                 n = 50000,
                                 include_rts = F,
                                 lang = "en",
                                 retryonratelimit = TRUE)
  
  ### Save the scraped data to a file that can be loaded later (for replicable analysis)
  saveRDS(debate_tweets, debate_tweets_rds)
  
}else{

  ### Load in data -------------------------------------------------
  debate_tweets <- readRDS(debate_tweets_rds)
}

names(debate_tweets)

debate_tweets %>% 
  head() %>% 
  View()

```


#### EDA

Which users tweeted about the debate the most?

```{r}

debate_tweets %>%
  count(screen_name, sort = T) %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = reorder(screen_name, n))) +
  geom_col()

```

What country did are people tweeting from?

```{r}

debate_tweets %>%
  count(country, sort = T) %>%
  filter(!is.na(country)) %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = reorder(country, n))) +
  geom_col() +
  scale_x_log10()

```

Which states in the US?

```{r}

debate_tweets %>%
  filter(location != "",
         country == "United States") %>%
  count(location, sort = T) %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = reorder(location, n))) +
  geom_col()

```


Common hashtags

**NOTE:** Hashtags are stored as a list in the column since users can have multiple hashtags per tweet.

```{r}

data.frame(hashtags = unlist(debate_tweets$hashtags))%>%
  count(hashtags, sort = T) %>%
  filter(hashtags != "Debate2020") %>% 
  top_n(25) %>%
  ggplot(aes(x = n, y = reorder(hashtags, n))) +
  geom_col()

```


#### Prepare Data for Sentiment Analysis

create a smaller data frame of only a few key columns

```{r}

debate_small <- debate_tweets %>%
  select(screen_name, country, location, text)

debate_small %>%
  head()

```


Get sentiment scores

**NOTE:** `afinn` is a sentiment lexicon that will score words between values of -5 to 5 based on whether they are negative or positive words. `bing` is a lexicon that categorizes words into either positive or negative categories

https://www.tidytextmining.com/

```{r}

tokens <- debate_small %>% 
  mutate(
    tweet_id = row_number()
  ) %>% 
  unnest_tokens(word, text)

token_sentiment <- tokens %>%
  inner_join(get_sentiments("afinn")) %>%
  inner_join(get_sentiments("bing"))

```

Histogram of scores

```{r}

token_sentiment %>%
  ggplot(aes(x = value)) +
  geom_histogram() + 
  scale_x_continuous(breaks = -5:5) +
  theme(
    panel.grid.minor.x = element_blank()
  )

```

The most positive words used

```{r}

token_sentiment %>%
  count(word, sentiment, sort = T) %>%
  filter(sentiment == "positive") %>%
  top_n(10) %>%
  ggplot(aes(x = n, y = reorder(word, n))) +
  geom_col()

```

The most negative words used

```{r}

token_sentiment %>%
  count(word, sentiment, sort = T) %>%
  filter(sentiment == "negative") %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = reorder(word, n))) +
  geom_col()

```


Average score by the top 4 country

```{r}

top_countries <- debate_tweets %>%
  count(country, sort = T) %>%
  filter(!is.na(country)) %>% 
  top_n(3) %>%
  pull(country)

token_sentiment %>%
  mutate(
    country = case_when(
      country %in% top_countries ~ country,
      TRUE ~ "Other")
  ) %>%
  group_by(country,tweet_id) %>%
  summarize(
    Avg_Sentiment_tweet = mean(value)
    ) %>%
  group_by(country) %>%
  summarize(
    mean_sentiment_country = mean(Avg_Sentiment_tweet),
    median_sentiment_country = median(Avg_Sentiment_tweet)
    ) %>% 
  pivot_longer(
    cols = c(mean_sentiment_country, median_sentiment_country),
    names_to = "stat",
    values_to = "value"
  ) %>% 
  ggplot( aes(x = value,y = country)) +
  geom_col(fill = "pale green",
           color = "black",
           alpha = 0.4) +
  geom_vline(aes(xintercept = 0),
             color = "red",
             size = 1.2) +
  facet_grid(~stat) +
  theme_classic() +
  theme(panel.grid.major.x = element_line(color = "white"),
        plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black"),
        axis.text = element_text(color = "white", 
                                 face = "bold",
                                 size = 12),
        axis.title = element_text(color = "white", 
                                 face = "bold",
                                 size = 15),
        plot.title = element_text(size = 20, 
                                  color = "white"),
        plot.subtitle = element_text(size = 17, 
                                     color = "white")) +
  xlim(-1, 1) +
  labs(x = "Average Sentiment",
       y = "",
       title = "Average #Debate2020 Sentiment",
       subtitle = "Top 3 Countries, and Other")

```



