---
title: "Tidyx Episode 77 - Make your models look clean with broom and tidymodels"
author: "Patrick Ward"
output: html_document
---

**tidymodels:** A collection of packages for modeling and machine learning within in the `tidyverse` framework.
**broom:** A package for cleaning up the *typically* messy outputs of models in `r`


## Packages & Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(broom)
library(Lahman)

theme_set(theme_light())

## Get baseball data
baseball <- Batting %>%
  inner_join(People) %>%
  inner_join(AwardsPlayers) %>%
  filter(yearID >= 2006,
         AB >= 300)

```


Always visualize your data!

```{r}

plot(baseball$H, baseball$RBI)

```

## Typical way of looking at the model output

```{r}
## Normal R method
old_fit <- lm(RBI ~ H, data = baseball)
summary(old_fit)
```

## Build a linear model in tidymodels

Initialize the linear model engine


```{r}

lm_mod <- linear_reg() %>% 
  set_mode("regression") %>% ## Not strictly necessary if we had a different model
  set_engine("lm") 

lm_mod

```


Fit the model

```{r}

fit_lm <- lm_mod %>%
  fit(RBI ~ H, data = baseball)

fit_lm
```

Summarize Model

```{r}
# model output from our tidymodel
fit_lm %>%
  pluck("fit") %>%
  summary()

# Get the same model output from the standard classic model
summary(old_fit)

```

**Let's use broom to tidy this up!**

```{r}
## model coefficients and standard errors
fit_lm %>%
  tidy()

## model fit statistics
fit_lm %>%
  glance()

## append fitted and residual values to the original data frame
fit_lm %>%
  pluck("fit") %>%
  augment()

## plot model coefficients with standard errors
fit_lm %>%
  tidy() %>%
  ggplot(aes(x = estimate, y = term)) +
  geom_point(size = 4) +
  geom_errorbar(aes(xmin = estimate - std.error,
                    xmax = estimate + std.error),
                width = 0,
                size = 1.2) +
  facet_wrap(~term, scales = "free") +
  expand_limits(x = 0) +
  theme(strip.background = element_rect(fill = "black"),
        strip.text = element_text(face = "bold"),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 12, face = "bold"))
```


## Making Predictions in tidymodels

create a fake data set

```{r}

set.seed(11)
baseball_data <- tibble(
  player = LETTERS[1:10],
  H = round(rnorm(n = 10, mean = 220, sd = 100), 0))

baseball_data
```


get point estimate prediction

```{r}
predict(fit_lm, 
        new_data = baseball_data)
```


get confidence intervals for predictions

```{r}
predict(fit_lm, 
        new_data = baseball_data, 
        type = "conf_int",
        level = 0.95)
```


combine predictions and confidence intervals with new data

```{r}
baseball_data_preds <- baseball_data %>%
  bind_cols(
    predict(fit_lm, 
        new_data = baseball_data)
  ) %>%
  bind_cols(
    predict(fit_lm, 
        new_data = baseball_data, 
        type = "conf_int",
        level = 0.95)
  )


baseball_data_preds

```


plot predictions 


```{r}
baseball_data_preds %>%
  rename(pred_rbi = .pred, low95 = .pred_lower, high95 = .pred_upper) %>%
  ggplot(aes(x = pred_rbi, y = reorder(player, pred_rbi))) +
  geom_errorbar(aes(xmin = low95, xmax = high95),
                width = 0,
                size = 1.2) +
  geom_point(size = 4,
             shape = 21,
             fill = "white",
             color = "black")
```

