---
title: "TidyX Episode 109 - R Classes - Making An S3 Object"
output: html_document
editor_options: 
  chunk_output_type: console
---

## R Objects

- Rich Object system (Base, S3, S4, R6, RC, ...R7?)
- Large % of R packages use S3

See Hadley Wickham's thoughts in Advanced R -
https://adv-r.hadley.nz/oo.html

## Benefits of an S3 Object

- Relatively simple to initialize
- Piggyback off of existing infrastructure
- Incredibly Flexible and easily expandable

- Initialize with `structure()` function

```{r}

my_s3 <- structure(
  .Data = 24
)

my_s3

```

## Defining your Object

- use "class" tag to add a _new_ class
- set various internal attributes using other tags
- Add data/non meta information via .Data

```{r}

my_s3 <- structure(
  .Data = list(
    message = "TidyX is Awesome!",
    authors = c("Ellis", "Patrick")
  ),
  creation_time = Sys.time(),
  class = "custom_message"
)

my_s3

```

## Developing Methods

- print, format, etc are all "generics", applying a method developed for the specific class
- ie. predict across all the model types!

```{r}

print

print.custom_message <- function(x,...){
  
  print_message <- c(
    "< Start Message >",
    paste0("  Contents: ",x$message),
    paste0("  Author: ", paste0(x$authors, collapse = ", ")),
    paste0("  Creation Time: ", format(attr(x, "creation_time"))),
    "< End Message >"
  )
  
  cat(
    print_message,
    sep = "\n"
  )
  
}

my_s3

```

## Develop Class Creator

- Avoid calling `structure()` directly
- Abstract some of the concepts needed

```{r}

new_message <- function(message, by){
  
  structure(
    .Data = list(
      message = message,
      authors = by
    ),
    creation_time = Sys.time(),
    class = "custom_message"
  )
  
}

m1 <- new_message(
  "You should Like and Subscribe to TidyX",
  by = "Ellis"
)
 m1

new_message(
  "Consider becoming a Patron at Patreon - patreon.com/Tidy_Explained",
  by = c("Patrick", "Ellis")
)

```


