---
title: "Episode 84 - Workflow sets"
author: "Patrick Ward"
output: html_document
---

## Load packages & data

**Insurance dataset from:** https://www.kaggle.com/mirichoi0218/insurance

```{r}

library(tidyverse)
library(tidymodels)

theme_set(theme_light())

insurance <- read_csv(here::here("TidyTuesday_Explained","084-Tidy_Models_8","insurance.csv"))

dim(insurance)
insurance %>% head()

insurance %>%
  summarize(across(.cols = everything(),
                   ~sum(is.na(.x))))

insurance %>%
  ggplot(aes(x = log(charges))) +
  geom_histogram()

```


**Predict the cost of insurance [charges] based on the 6 features**


## Split data

```{r}

## Train/Test
set.seed(8675309)
insurance_split <- initial_split(insurance, strat = "region")

train <- training(insurance_split)
test <- testing(insurance_split)

## Cross Validation Split of Training Data
set.seed(33)
cv_folds <- vfold_cv(
  data = train, 
  v = 5
  ) 

cv_folds

```


## Recipe

```{r}

insurance_rec <- recipe(
    charges ~ .,
    data = train
  ) %>%
  step_dummy(region, one_hot = TRUE) %>%
  step_log(charges)

insurance_rec %>%
  prep() %>%
  bake(new_data = NULL)

```


## Model Specifications

1) Linear Regression
2) Random Forest
3) KNN Regression


```{r}

lm_spec <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

rf_spec <- rand_forest() %>%
  set_mode("regression") %>%
  set_engine("randomForest", importance = TRUE)

knn_spec <- nearest_neighbor(neighbors = 4) %>%
  set_mode("regression")

```


## Workflow Set

* Combine the pre-processing recipe and the three models together

```{r}


wf_set <-workflow_set(
  preproc = list(insurance_rec),
  models = list(lm_spec, rf_spec, knn_spec)
  )

wf_set

```


## Fit the models to the workflow

```{r}

doParallel::registerDoParallel()

insurance_fit <- workflow_map(
    wf_set,
    "fit_resamples",
    resamples = cv_folds,
    seed = 22 ## replicability
  )

insurance_fit

```


## Evaluate model fits

```{r}
autoplot(insurance_fit)

collect_metrics(insurance_fit)

rank_results(insurance_fit, rank_metric = "rmse", select_best = TRUE)
```


## Extract the best workflow

```{r}

wf_final <- extract_workflow(insurance_fit, id = "recipe_rand_forest") %>%
  fit(train)

wf_final

library(vip)

wf_final %>%
  extract_fit_parsnip() %>%
  vip(geom = "col")

```



## Predict on test set


```{r}

preds <- predict(
  wf_final %>% extract_fit_parsnip(), 
  new_data = insurance_rec %>% prep() %>% bake(new_data = test)
)

preds %>% head()

test_final <- cbind(test, preds)
test_final %>% head()


raw_plt <- test_final %>%
  ggplot(aes(x = exp(.pred), y = charges)) +
  geom_point(aes(color = region),
             size = 2,
             alpha = 0.5
             ) +
  geom_smooth(method = "lm",
              color = "red",
              size = 2) +
  geom_abline(intercept = 0,
              slope = 1,
              size = 2,
              color = "green",
              linetype = "dashed") +
  ggtitle("Charges")

raw_plt

```

```{r}

## all in one
wf_final_full <- extract_workflow(insurance_fit, "recipe_rand_forest") %>%
  last_fit(insurance_split)

wf_final_full %>% collect_predictions()

```

