---
title: "R Classes - Dates"
author: "Ellis Hughes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Objects

- Rich Object system (Base, S3, S4, R6, RC, ...R7?)
- Define behavior through different semantics
- Large % of R packages use S3
- Bioconductor favors S4
- R6 behaves like a Python Class.

See Hadley Wickham's thoughts in Advanced R -
https://adv-r.hadley.nz/oo.html

## Date Type Objects (S3)

- Date
- POSIXct
- POSIXlt

### Date Objects

```{r Date-Objects}

# Provide an ISO date, easily converts
date1 <- as.Date("2022-01-01")

# a date object contains a number only
# number represents number of days since the origin 1970-01-01
dput(date1)

## Alternatively provide a number and an origin. 
### Number represents days
### Origin is date to start counting from
date2 <- as.Date(0, origin = "2022-01-01")
date3 <- as.Date(365, origin = "2021-01-01")

as.numeric(date1)
as.numeric(date2)
as.numeric(date3)

## Base R origin is 1970-01-01
as.numeric(as.Date(0, origin = "1970-01-01"))

## When converting dates in Excel, origin is 1899-12-31


```

## POSIXct

- POSIXct represents number of seconds from origin (1970-01-01)
- Sub-day units (down to milliseconds)
- Time Zones
- Multiple formats
  
```{r cars}

date_time1 <- as.POSIXct("2022-01-01 8:00:01 PM PST") 

date_time1
as.numeric(date_time1)
dput(date_time1)

as.numeric(date_time1)/60

## Alternatively provide a number and an origin. 
### Number represents seconds
### Origin is date to start counting from
### Gets confusing with timezone manipulation
date_time2 <- as.POSIXct(0, origin = "2022-01-01 00:00:00AM")
date_time3 <- as.POSIXct(365*24*60*60, origin = "2021-01-01 00:00:00AM", tz = "UTC")

```


## Working with Dates/Date-time

Lubridate Cheat Sheet: https://rawgit.com/rstudio/cheatsheets/main/lubridate.pdf

```{r pressure, echo=FALSE}

library(lubridate)

### parsing helpers - Lubridate getting started page 

?strptime

## 20110604
as.Date("20110604")
as.Date("20110604", format = "%Y%m%d")

ymd("20110604")
ymd("2011/06/04")
ymd("2011-06-04")
ymd("2011 06 04")

mdy("06042011")
dmy("04062011")


## 20110604 103000
as.POSIXct("20110604 103000", format = "%Y%m%d %H%M%S", tz = "UTC")
ymd_hms("20110604 103000")


## Time differences/Intervals
date1 <- as.POSIXct(0, origin = "2022-01-01")
date2 <- as.POSIXct(10, origin = "2022-01-01")
date3 <- as.POSIXct(10, origin = "2021-01-01")

### Base R
date2 - date1
date3 - date1

difftime(date2, date1, units = "mins")
10/60

difftime(date2, date1, units = "days")
difftime(date2, date1, units = "secs")

as.numeric(difftime(date2, date1, units = "secs"))
as.numeric(date2-date1)
as.numeric(date3-date1)


### Lubridate
d1d2_interval <- interval(date2,date1)
as.duration(d1d2_interval)

## Adding/subtracting time

### Base R
date1 <- as.POSIXct("2022-01-01")
date1 + 7 ## adds 7 seconds
as.Date(date1) + 7 ## adds 7 days

date1 + (7*24*60*60) ## to move forward 7 days in posixct

### Lubridate
date1 + dweeks(1)
as.Date(date1) + dweeks(1)

# dpicoseconds, ... dseconds, dminutes, dmonths, dyears

## Formatting outputs

### Base R
?strptime
format(date1, "%Y-%d-%m %H:%M %Z")
format(date1, "%y-%d-%m %H:%M %Z") ## case matters

```



