---
title: "Random Forests"
author: "Ellis Hughes & Patrick Ward"
date: "4/4/2021"
output: html_document
---

## References

**Data courtesy of {mlbgameday}:** https://github.com/keberwein/mlbgameday
**Resource for understanding pitchf/x data:** https://library.fangraphs.com/misc/pitch-fx/
**Resource on the features in the pitchf/x data:** Kagan, David. (2008). Fun with PitchFX Data. 

## Problem Statement

The FBL (fantasy baseball league) wants to predict which pitches are thrown. 
Traditionally it was labeled by a human sitting in the stands.
They want to productionalize it to give the classifications faster. 

They hired us...

## Load Data

```{r setup, include=FALSE}

bp <- here::here("TidyTuesday_Explained/055-MLB_pitch_classification_3")

knitr::opts_chunk$set(echo = TRUE,
                      root.dir = bp)

suppressPackageStartupMessages({
  suppressWarnings({
    library(tidyverse)
    library(rpart)
    library(rpart.plot)
    library(randomForest)
    library(caret)
    library(doParallel)
  })
})

theme_set(theme_light())

```

```{r load-data-and-clean}

## see Episode 53 - bit.ly/TidyX_Ep53 - for the background on this cleanup

# train
set1 <- readRDS(file.path(bp,"2016_04_21_to_2016_04_23_pitch.rds"))
set2 <- readRDS(file.path(bp,"2016_04_24_to_2016_04_26_pitch.rds"))
train <- bind_rows(set1, set2)

#test
test <- readRDS(file.path(bp,"2016_04_27_to_2016_04_30_pitch.rds"))

names(train)


#### Cleaning data based on EDA ----------------------------------------------

# for more details on decisions, Check out TidyX 53
train_cleaned <- train %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

test_cleaned <- test %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO", "SC"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

train_cleaned$pitch_type <- factor(train_cleaned$pitch_type, level = unique(train_cleaned$pitch_type))

test_cleaned$pitch_type <- factor(test_cleaned$pitch_type, level = levels(train_cleaned$pitch_type))


```


## Decision Trees

```{r decision-trees}

fit_tree <- rpart::rpart(pitch_type ~ ., data = train_cleaned)

summary(fit_tree)

plot(fit_tree)
text(fit_tree)

rpart.plot(fit_tree)

test_tree_pred <- bind_cols(
  test_cleaned,
  pred = predict(fit_tree, newdata = test_cleaned, type = "class"))

test_tree_pred %>%
  count(pred, pitch_type) %>%
  group_by(pitch_type) %>%
  summarize(N = sum(n),
            matches = sum(ifelse(pitch_type == pred, n, 0)),
            within_acc = matches / N)

# Overall Accuracy
confusionMatrix(test_tree_pred$pitch_type,test_tree_pred$pred)

sum(diag(table(test_tree_pred$pred, test_tree_pred$pitch_type))) / sum(table(test_tree_pred$pred, test_tree_pred$pitch_type))

```

## Random Forests

```{r random-forest}

set.seed(567)
fit_rf <- randomForest(pitch_type ~ ., data = train_cleaned, mtry = 6)
fit_rf

names(fit_rf)
fit_rf$importance
varImpPlot(fit_rf)

test_rf_pred <- bind_cols(test_cleaned, rf_pred = predict(fit_rf, newdata = test_cleaned, type = "class"))
test_rf_pred %>% head()

test_rf_pred %>%
  janitor::tabyl(rf_pred, pitch_type)

test_rf_pred %>%
  count(rf_pred, pitch_type) %>%
  group_by(pitch_type) %>%
  summarize(N = sum(n),
            matches = sum(ifelse(pitch_type == rf_pred, n, 0)),
            within_acc = matches / N)

# Overall Accuracy
confusionMatrix(test_rf_pred$pitch_type,test_rf_pred$rf_pred)


sum(diag(table(test_rf_pred$rf_pred, test_rf_pred$pitch_type))) / sum(table(test_rf_pred$rf_pred, test_rf_pred$pitch_type))

```


```{r random-forest-tuned-parallel}

## Tune the mtry parameter using the `caret` package
control <- trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 1,
  search = "grid")

set.seed(567)

## which param to try to tune on
tunegrid <- expand.grid(.mtry=c(3:7))

## This process can take a while, so lets make it faster with parallelization
cl <- makePSOCKcluster(5)
registerDoParallel(cl)
  
fit_rf_grid <- train(pitch_type ~ .,
                     data = train_cleaned,
                     method="rf",
                     metric = "Accuracy",
                     tuneGrid = tunegrid,
                     trControl = control)

## turn off the parallelization to close the R processes
stopCluster(cl)

fit_rf_grid

## Save the model
saveRDS(fit_rf_grid, file = file.path(bp,"optimal_rf.rds"))

## load model
fit_rf_grid <- readRDS(file.path(bp,"optimal_rf.rds"))

fit_rf_grid

fit_rf_grid$results %>% arrange(desc(Accuracy))

```


```{r visualize}
# plot
plot(fit_rf_grid)

## predict on test set

test_rf_pred2 <- bind_cols(test_cleaned, rf_pred = predict(fit_rf_grid, newdata = test_cleaned, type = "raw"))

test_rf_pred2 %>% head()

test_rf_pred2 %>%
  janitor::tabyl(rf_pred, pitch_type)

test_rf_pred2 %>%
  count(rf_pred, pitch_type) %>%
  group_by(pitch_type) %>%
  summarize(N = sum(n),
            matches = sum(ifelse(pitch_type == rf_pred, n, 0)),
            within_acc = matches / N)

# Overall Accuracy
confusionMatrix(test_rf_pred2$pitch_type,test_rf_pred2$rf_pred)

sum(diag(table(test_rf_pred2$rf_pred, test_rf_pred2$pitch_type))) / sum(table(test_rf_pred2$rf_pred, test_rf_pred2$pitch_type))

```
