---
title: 'Episode 126: pivot wider but keep duplicates'
author: "Patrick Ward"
output: html_document
---

`pivot_wider()` is a useful function for pivoting data from a long data set to a wide data set. The issue is that if the data has duplicated rows, `pivot_wider()` will default to converting the data to nested lists.

We will use an example from the `mtcars` data set using the `cyl` and `wt` columns and we will start by converting `cyl` to a factor variable since it only has three levels (4, 6, and 8).

```{r}
library(tidyverse)

dat <- mtcars %>%
  select(wt, cyl) %>%
  mutate(cyl = as.factor(cyl))

dat %>%
  head()
```


**Goal:** We want to pivot the data so that `wt` is the id column and `cyl` to be pivoted wider.

```{r}

tribble(
   ~ wt, ~`8`,  ~`6`, ~`4`,
  2.620,    0,     1,    0, 
  2.875,    0,     1,    0, 
  2.320,    0,     0,    1, 
  2.320,    0,     0,    1, 
  3.215,    0,     1,    0, 
  3.440,    1,     0,    0, 
  3.440,    0,     1,    0, 
  3.460,    0,     1,    0
)

```

## First Attempt (and failure)

```{r}
dat %>%
  pivot_wider(
    values_from = cyl, 
    names_from = cyl
  )

```

* We get an issue about the fact that there are duplicates with respect to `wt` and `cyl`, which causes R to put the results into a list and gives us a warning with some ways to solve the issue.

## Second Attemt (and failure)

```{r}
dat %>%
  pivot_wider(values_from = cyl, 
              names_from = cyl,
              values_fn = length)
```

* At first glace, this looks like adding the `values_fn = length` argument has worked! But notice that this is a tible of 29 rows. The original data has 32 rows.

```{r}
nrow(dat)
```


* Notice that if there is a duplicate, the `values_fn = length` function adds a `2` under the respecitve `cyl` column (e.g., row 5).

## Third Approach (and fail)

```{r}
dat %>%
  group_by(wt, cyl) %>%
  mutate(wt_id = row_number()) %>% 
  pivot_wider(values_from = cyl, 
              names_from = cyl,
              id_cols = c(wt_id, wt),
              values_fn = length)
```


* So close!! We have 31 rows instead of 32, though.
* The issue is `wt == 3.44`

```{r}
dat %>%
  select(cyl, wt) %>% 
  group_by(wt, cyl) %>%
  mutate(wt_id = row_number()) %>%
  filter(wt == 3.44)
```


## Fourth Approach (success!)

* Creat an id variable for `wt`
* Create an id variable for `wt` and `cyl` together
* Also, get rid of the NA's and put the prefix `cyl` in front of the pivoted columns

```{r}
dat %>%
  select(cyl, wt) %>% 
  group_by(wt, cyl) %>%
  mutate(wt_id = row_number(),
         wt_cyl_id = paste(wt_id, cyl, sep = "_")) %>% 
  pivot_wider(values_from = cyl, 
              names_from = cyl,
              id_cols = c(wt_cyl_id, wt),
              values_fill = 0,
              names_prefix = "cyl",
              values_fn = length)
```

