---
title: "K Nearest Neighbors and UMAP"
author: "Ellis Hughes & Patrick Ward"
date: "3/20/2021"
output: html_document
---

## References

**Data courtesy of {mlbgameday}:** https://github.com/keberwein/mlbgameday
**Resource for understanding pitchf/x data:** https://library.fangraphs.com/misc/pitch-fx/
**Resource on the features in the pitchf/x data:** Kagan, David. (2008). Fun with PitchFX Data. 

## Problem Statement

The FBL (fantasy baseball league) wants to predict which pitches are thrown. 
Traditionally it was labeled by a human sitting in the stands.
They want to productionalize it to give the classifications faster. 

They hired us...

## Load Data

```{r setup, include=FALSE}

bp <- here::here("TidyTuesday_Explained/054-MLB_pitch_classification_2")

knitr::opts_chunk$set(echo = TRUE,
                      root.dir = bp)

suppressPackageStartupMessages({
  suppressWarnings({
    library(tidyverse)
    library(plotly)
    library(class)
    library(caret)
    library(umap)
  })
})

theme_set(theme_light())

# train
set1 <- readRDS(file.path(bp,"2016_04_21_to_2016_04_23_pitch.rds"))
set2 <- readRDS(file.path(bp,"2016_04_24_to_2016_04_26_pitch.rds"))
train <- bind_rows(set1, set2)

#test
test <- readRDS(file.path(bp,"2016_04_27_to_2016_04_30_pitch.rds"))

names(train)


#### Cleaning data based on EDA ----------------------------------------------

# for more details on decisions, Check out TidyX 53
train_cleaned <- train %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

test_cleaned <- test %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO", "SC"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)
```

## KNN

```{r}

scale_by <- function(x, mean, sd){
  (x - mean) / sd
}

scale_values <- train_cleaned %>% 
  select(-pitch_type) %>% 
  lapply(function(x){
    list(mean = mean(x), sd = sd(x))
  })

## prepare the training data
train_pitches <- train_cleaned$pitch_type

train_cleaned_scaled <- colnames(train_cleaned) %>% 
  setdiff("pitch_type") %>%
  lapply(function(column_name){
    scale_by(
      x = train_cleaned[[column_name]],
      mean = scale_values[[column_name]][["mean"]],
      sd = scale_values[[column_name]][["sd"]]
    )
  }) %>% 
  do.call('data.frame',.) %>% 
  setNames(colnames(train_cleaned) %>%  setdiff("pitch_type"))

## prepare the testing data
test_pitches <- test_cleaned$pitch_type
test_cleaned_scaled <- colnames(test_cleaned) %>% 
  setdiff("pitch_type") %>%
  lapply(function(column_name){
    scale_by(
      test_cleaned[[column_name]],
      scale_values[[column_name]][["mean"]],
      scale_values[[column_name]][["sd"]]
    )
  }) %>% 
  do.call('data.frame',.) %>% 
  setNames(colnames(test_cleaned) %>% setdiff("pitch_type"))

## run a KNN
fit_knn <- knn(
  train = train_cleaned_scaled,
  test = test_cleaned_scaled,
  cl = train_pitches,
  k = 4, ## based on last weeks observations
  prob = TRUE
)

## create a table of predicted and observed outcomes for the test set
pred <- data.frame(Pred = fit_knn, Observed = test_pitches)
table("Predicted" = pred$Pred, "Observed" = pred$Observed)

## calculate the misclassification rate
##  Overall
miss_classification_rate <- 1 - sum(pred$Pred == pred$Observed) / nrow(test_cleaned_scaled)
miss_classification_rate

## within pitch type
table("Predicted" = pred$Pred, "Observed" = pred$Observed) %>%
  data.frame %>% 
  group_by(Observed) %>%
  summarise( mcr = 1 - Freq[Observed == Predicted] / sum(Freq))

## get the probability of the majority vote
pred_tbl <- bind_cols(pred, prob_class = attr(fit_knn, "prob"))
pred_tbl %>% head()

```

## optimizing

```{r optimizing-knn}

tc <- bind_cols(train_cleaned_scaled, train_pitches = train_pitches)
tc2 <- bind_cols(test_cleaned_scaled, train_pitches = test_pitches)


set.seed(555)
ctrl <- trainControl(method="repeatedcv",repeats = 1) #normally use 5 for 5 fold 
knnFit <- train(
  train_pitches ~ ., 
  data = tc, 
  method = "knn",
  trControl = ctrl)

#Output of kNN fit
knnFit

plot(knnFit)


knnPredict <- predict(knnFit,newdata = tc2)

table(tc2$train_pitches, knnPredict)


## within pitch type
table("Predicted" = knnPredict, "Observed" = tc2$train_pitches) %>% data.frame %>% group_by(Observed) %>% summarise( mcr = 1- Freq[Observed == Predicted] / sum(Freq), n = sum(Freq))

```

```{r knn-comparison}

k_4 <- table("Predicted" = pred$Pred, "Observed" = pred$Observed) %>%
  data.frame %>% 
  group_by(Observed) %>%
  summarise( mcr = 1 - Freq[Observed == Predicted] / sum(Freq))

table("Predicted" = knnPredict, "Observed" = tc2$train_pitches) %>%
  data.frame %>% 
  group_by(Observed) %>% 
  summarise(
    mcr = 1 - Freq[Observed == Predicted] / sum(Freq),
    n = sum(Freq)) %>% 
  left_join(k_4, by = "Observed", suffix = c("_kOpt", "_k4")) %>% 
  relocate(n,.after = ends_with("_K4"))

```


```{r UMAP}


umap_pitch <- train_cleaned %>% 
  select(-pitch_type) %>% 
  as.matrix %>% 
  umap()

## run previously
# umap_pitch_50 <- train_cleaned %>% 
#   select(-pitch_type) %>% 
#   as.matrix %>% 
#   umap(
#     n_neighbors = 50
#   )
# saveRDS(umap_pitch_50,file.path(bp,"UMAP_50.rds"))

# umap_pitch_50 <- readRDS(file.path(bp,"UMAP_50.rds"))

pitch_vect <- c(
  FF = "four-seam fastball",
  SL = "slider",
  CH = "change up",
  FT = "two-seam fastball",
  SI = "sinker",
  CU = "curveball",
  FC = "cutter",
  KC = "knuckle-curve",
  FS = "split-finger fastball"
)




pitch_vect[train_cleaned$pitch_type]

umap_plot <- umap_pitch %>% 
  pluck("layout") %>% 
  data.frame %>% 
  setNames(c("umap1","umap2")) %>% 
  bind_cols(umap_pitch$data) %>% 
  mutate(
    pitch_type = pitch_vect[train_cleaned$pitch_type]
  ) %>% 
  ggplot(aes(x = umap1, y = umap2)) + 
  geom_point(aes( color = pitch_type), alpha = .5)

ggplotly(umap_plot)

# umap_plot_50 <- umap_pitch_50 %>% 
#   pluck("layout") %>% 
#   data.frame %>% 
#   setNames(c("umap1","umap2")) %>% 
#   bind_cols(umap_pitch$data) %>% 
#   mutate(
#     pitch_type = pitch_vect[train_cleaned$pitch_type]
#   ) %>% 
#   ggplot(aes(x = umap1, y = umap2)) + 
#   geom_point(aes( color = pitch_type), alpha = .5)
# 
# ggplotly(umap_plot_50)

```