---
title: "TidyX 19"
author: "Patrick Ward"
date: "7/14/2020"
output: html_output
editor_options: 
  chunk_output_type: console
---

## Load Packages & Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, warning = FALSE)

## packages & functions

suppressPackageStartupMessages({suppressWarnings({
  library(tidyverse)
  library(tidytuesdayR)
  library(plotly)
  library(formattable)
  library(sparkline)
})})


theme_set(theme_light())

pct_total <- function(x){
  p = x / sum(x)
  return(p)
}

```

```{r load-data}
## load data

tt_data <- tidytuesdayR::tt_load('2020-07-14')

space <- tt_data$astronaut

```


## EDA

**Look at the data**

```{r peek at the data}

glimpse(space)
colnames(space)
dim(space)

## Any missing data?
space %>%
  summarize(across(everything(), ~sum(is.na(.x)))) %>%
  t()

```


**Simple Visuals**

**Note: I make an assumption here that `nationality` of the astronaut represents the country that they traveled to space for. This may or may not be the case.*


How many unique countries are there?

```{r}

space %>%
  distinct(., nationality) %>%
  nrow() %>% 
  paste("unique countries", sep = " ")

```


Which countries have gone to space most often?

```{r}
space %>%
  count(nationality, sort = T) %>%
  head() %>%
  ggplot(aes(x = n, y = reorder(nationality, n))) +
  geom_col() +
  geom_label(aes(label = n))
```


What years have space missions taken place?

```{r}
space %>%
  count(year_of_mission, sort = T) %>%
  ggplot(aes(x = year_of_mission, y = n)) +
  geom_col(fill = "blue",
           alpha = 0.6,
           color = "black")
```


What has been the space mission trajectory across decades for the top 6 countries to have space missions?

##### Build a plotly figure for this

```{r}
## still plot
space %>%
  filter(nationality %in% c("U.S.",
                            "U.S.S.R/Russia",
                            "Japan",
                            "France",
                            "Canada",
                            "Germany")) %>%
  mutate(Decade = floor(year_of_mission / 10) * 10) %>%
  group_by(nationality, 
           Decade) %>%
  summarize(Number_of_Missions = max(mission_number)) %>%
  ggplot(aes(x = Decade, 
             y = Number_of_Missions,
             color = nationality)) +
  geom_line(size = 1.1)

################################################################
### plotly -----------------------------------------------------
## Create a highlighter for the countries
highlight_country <- space %>%
  filter(nationality %in% c("U.S.",
                            "U.S.S.R/Russia",
                            "Japan",
                            "France",
                            "Canada",
                            "Germany")) %>%
  mutate(Decade = floor(year_of_mission / 10) * 10) %>%
  group_by(nationality, 
           Decade) %>%
  summarize(Number_of_Missions = max(mission_number)) %>%
  highlight_key(., ~nationality)

# initiate the plotly object
fig <- plot_ly(
  highlight_country,
  color = I("black")) %>%
  group_by(nationality)

# create the time series plot
time_series <- fig %>%
  group_by(nationality) %>%
  add_lines(x = ~Decade,
            y = ~Number_of_Missions)

# Create interactive plot with text box to select country of interest
highlight(
  time_series,
  on = "plotly_click",
  selectize = TRUE,
  dynamic = TRUE,
  persistent = TRUE)

```


How many total missions have the different astronauts performed?

```{r}

space %>%
  count(total_number_of_missions, sort = T) %>%
  ggplot(aes(x = n, y = reorder(total_number_of_missions, n))) +
  geom_col()

```


Average number of missions per country?

```{r}

space %>%
  distinct( name, .keep_all = T) %>%
  group_by(nationality) %>%
  summarize(Avg_Num_Missions = mean(total_number_of_missions)) %>%
  arrange(desc(Avg_Num_Missions))

```


How many male and female astronauts have there been?

```{r}
space %>%
  count(sex) %>%
  ggplot(aes(x = sex, y = n)) +
  geom_col()
```


## Building a Dashboard

**Variables to include**

* Nationality (country)
* Number of missions per decade
* Number of different astronauts who have had a mission
* Percentage of female astronauts who have had missions

Create data frames

```{r}

## Astronauts per decade data frame
decades <- space %>%
  mutate(Decade = floor(year_of_mission / 10) * 10) %>%
  group_by(nationality, 
           Decade) %>%
  summarize(Number_of_Astronauts = length(unique(name))) %>%
  arrange(Decade) %>%
  pivot_wider(data = .,
              names_from = Decade,
              values_from = Number_of_Astronauts,
              values_fill = 0)

## Count of the number of unique astronauts per country
astronauts <- space %>%
  group_by(nationality) %>%
  distinct(., name) %>%
  summarize(N_Astronauts = n())

## Percent Female Astronauts
female_astronauts <- space %>%
  group_by(nationality, sex) %>%
  distinct(., name) %>%
  group_by(nationality) %>%
  count(sex) %>%
  mutate(Pct_Female = n / sum(n)) %>%
  filter(sex == "female") %>%
  select(nationality, Pct_Female)

## Join tables
df_final <- decades %>%
  left_join(astronauts) %>%
  left_join(female_astronauts) %>%
  mutate(Pct_Female = ifelse(is.na(Pct_Female), 0, Pct_Female))

df_final %>% head()
```


Build Dashboard

```{r}

## create sparkline line 
df_final$`Trend Line` <- 
  apply(df_final[, 2:7], 
        MARGIN = 1, 
        FUN = function(x){
          as.character(
            htmltools::as.tags(
              sparkline(as.numeric(x), type = "line")
              ))
          })


## Create plot as an html widget
output_dashboard <- df_final %>%
  rename(
    Nationality = nationality,
    `Total Astronauts` = N_Astronauts,
    `% Female` = Pct_Female
  ) %>% 
  arrange(desc(`Total Astronauts`)) %>%
  formattable(
    align = c("l", rep("c", NCOL(df_final) - 4), rep("r", 2)), # Left align the 
    list(
      `Total Astronauts` = color_bar("#CCFF33", fun = pct_total),
      `% Female` = percent)
    ) %>% 
  as.htmlwidget()

# add sparkline dependencies
output_dashboard$dependencies <- c(
  output_dashboard$dependencies, 
  htmlwidgets::getDependency("sparkline", "sparkline"))

# run dashboard
output_dashboard

# save dashboard
htmlwidgets::saveWidget(
  widget = output_dashboard,
  title = "Seven Decades of Space",
  file = here::here(
    "TidyTuesday_Explained/019-Astronauts_and_dashboards/Seven Decades of Space.html"
    ),
  selfcontained = TRUE
)
```

