---
title: "Creating Subject ID's and Withing Subject Observation Counters"
output: html_document
---

## Data

```{r}
library(tidyverse)

dat <- tibble(
  Subject = as.factor(c(rep("Al", 20), rep("Tom", 35), rep("Kerry", 13))),
  variable = round(rnorm(mean = 300, sd = 50, n = Subject), 0)
  ) %>% 
  group_by(Subject) %>% 
  mutate(
    observation = seq_along(Subject)
  ) %>% 
  ungroup()

dat
```


## How many observations per subject

```{r}

dat %>%
  count(Subject, sort = TRUE)

```



## Unique IDs

Give each subject an id (in case we don't want to expose their names in our study)

```{r}

dat_id <- dat %>%
  group_by(Subject) %>%
  mutate(index = cur_group_id())





## random order
rand_dat <- dat %>% 
  .[sample(1:nrow(dat), nrow(dat), replace = FALSE),] %>%
  group_by(Subject) %>%
  mutate(index = cur_group_id()) 





## joining
dat_join <- tibble(Subject = c("Al", "Kerry", "Tom", "Jerry"),
                   index = c(1, 2, 3, 4)) %>%
  right_join(dat, by = "Subject")



dat_id %>% 
  arrange(index, observation) %>% 
  head()

rand_dat %>% 
  arrange(index, observation) %>% 
  head()

dat_join %>% 
  arrange(index, observation) %>% 
  head()

dat_id %>%
  count(Subject, index)

```


## Creat an ID for groupings of 5 observations

What if we want to break up each subject's observations into groupings of 5? Also, some of the subjects do not have a total number of observations divisble by 5, so we need to make sure that we count by 5's but can end on a non-five divisible number.

To do this, we use integer division, which simply divides two numbers, discards the remainder and leaves the integer.

For example, 10/7 = 1.4. The result for integer vision will be 1, discarding the .4

```{r}

10/7
10 %/% 7

```

Let's try this on one subject before applying it to everyone.

* Kerry only had 13 observations. Pull just her observations

```{r}

kerry <- dat_id %>% 
  filter(Subject == "Kerry") %>% 
  pull(variable)

```

* We can count along an index of each observation using the `seq_along()` function

```{r}

seq_along(kerry)

```

* Let's apply integer division to each of the indexes in the sequence

```{r}

seq_along(kerry) %/% 5

```

* Integer division is giving us the groupings in 4's
* What if we subtract each index number by 1, first?

```{r}

(seq_along(kerry) - 1) %/% 5

```


* That works! Now we just need to add 1 to the results and we have a counter by 5's

```{r}
(seq_along(kerry) - 1) %/% 5 + 1
```

* Now apply it to all subjects

```{r}

dat_id_grouped <- dat_id %>%
		group_by(Subject) %>%
		mutate(group5 = (seq_along(Subject) - 1) %/% 5 + 1)

dat_id_grouped %>%
  knitr::kable(align = 'c')

```

