---
title: "TidyX Episode 96 - Rmarkdown Guide - ASIS results"
author: "TidyX"
output:
  html_document:
    df_print: paged
---

## R Chunk Options Review

There is an opportunity to control a lot of behavior of code execution and results

- title
- warnings/messages
- image dimensions
- code display

```{r this-is-the-title, warning=FALSE, message=FALSE}

cat(c(1:5))

```


## Resuts Options

4 Options exist for "results" chunk options
  - markup
  - asis
  - hold
  - hide
  
Their behaviors impact how the results are displayed at a highest level:

  - markup
    - default option, adds markup to outputs
    
  - asis
    - results are added to markdown document "asis"
    
  - hold
    - "holds" outputs of code until entire chunk has been evaluated
    
  - hide
    - no results are displayed

## ASIS

Lots of interesting impact potential to report, mainly potential for creation of dynamic content!

### default
```{r, echo = FALSE}

cat(" This is some great content!\n\n")
cat(" If you like TidyX, be sure to [subscribe!](https://www.youtube.com/tidyX_screencast?sub_confirmation=1)")

```


### asis
```{r, results = "asis", echo = FALSE}

cat(" This is some great content!\n\n")

cat(" If you like TidyX, be sure to [subscribe!](https://www.youtube.com/tidyX_screencast?sub_confirmation=1)")

```


## pandoc under the hood

rmarkdown converts your document into a markdown document that is then converted into the output set (html, word, pdf) through pandoc.
pandoc is a cool piece of software included in your rstudio install.

Using this knowledge, pandoc allows you to pass special code in your rmarkdown depending out the output set.

<h3> Since this output is HTML, I can write <span style = "font-style:italic; color:blue;">inline</span> html </h3>

This also applies to as-is outputs

```{r, results = "asis", echo = FALSE}


cat("<h3> THIS IS AN  <span style = 'font-style:italic; color:blue;'>ASIS</span> OUTPUT </h3>")

```
