---
title: "Wine_Word_Cloud"
author: "Ellis Hughes"
date: "3/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidytuesdayR)
library(tidytext)
library(tidyverse)
library(reshape2)
library(wordcloud)

```

## Wine Mag Ratings

This dataset comes from Wine Enthusiast, scrapped from their website.
It was posted to Kaggle for folks to have fun going through the data, and was scraped November 2017

```{r load_data}
wine_mag <- tt_load(2019,week = 22)
wine_mag
```

## Description Sentiment vs Score

My curiousity is based on the sentiment of the description, is there on average a higher score?
I would assume there would be, but wine might have some unique word meanings not identified by the standard bing sentiment.

```{r parse-descriptions}
wine_mag_tokened <- wine_mag$`winemag-data-130k-v2` %>% 
  unnest_tokens(output = word,input = description) %>% 
  anti_join(get_stopwords(), by = "word") %>% 
  glimpse()

```

Get sentiments based on the "Bing" data for each word.

```{r bing-sentiment}

bing <- get_sentiments("bing")

wine_mag_bing_sentiments <- wine_mag_tokened %>% 
  inner_join(bing, by = "word")

```

Take stock of what the top words.

```{r top-words}

wine_mag_bing_sentiments %>%
  group_by(word) %>% 
  summarise(n = n(), sentiment = first(sentiment)) %>%
  arrange(desc(n)) %>% 
  head(n=10)
  
```

Calculate Wine sentiment by score:

```{r score-sentiment}

wine_mag_score_sentiments <- 
  wine_mag_bing_sentiments %>%
  group_by(points) %>%
  summarize(
    n = n(),
    negative = sum(sentiment == "negative"),
    positive = sum(sentiment == "positive")
  ) %>%
  mutate(norm_neg = negative / n,
         norm_pos = positive / n) %>%
  mutate(sentiment_score = norm_pos - norm_neg)

wine_mag_score_sentiments %>% 
  mutate(sent_dummy = ifelse(sentiment_score<0, 'More Negative', 'More Positive')) %>%
  mutate(points = factor(points)) %>%
  ggplot() +
  geom_bar(
    aes(
      x = points,
      y = sentiment_score,
      fill = sent_dummy
    ),
    stat = "identity"
  ) + 
  coord_flip() +
  labs(y='Average sentiment of scores \n (Positive - Negative Words)/(total words)', x = 'Score') +
  theme_bw() + 
  theme(legend.position='none', axis.text.x=element_text(size=12), axis.title.x = element_text(size=14, face = 'bold'),
        axis.text.y=element_text(size=12), axis.title.y=element_text(size=14, face='bold'))

```

## Obligatory Word Cloud

Generate word cloud of most positive and negative words

```{r word-cloud, echo = FALSE}


whites <-
  c(
    "Albariño",
    "Bordeaux-style White Blend",
    "Champagne Blend",
    "Chardonnay",
    "Chenin Blanc",
    "Gewürztraminer",
    "Glera",
    "Grüner Veltliner",
    "Moscato",
    "Pinot Blanc",
    "Pinot Grigio",
    "Pinot Gris",
    "Portuguese White",
    "Rhône-style White Blend",
    "Riesling",
    "Sauvignon Blanc",
    "Sparkling Blend",
    "Viognier",
    "White Blend"
  )
reds <-
  c(
    "Aglianico",
    "Barbera",
    "Bordeaux-style Red Blend",
    "Cabernet Franc",
    "Cabernet Sauvignon",
    "Carmenère",
    "Corvina, Rondinella, Molinara",
    "Gamay",
    "Garnacha",
    "Grenache",
    "Malbec",
    "Merlot",
    "Nebbiolo",
    "Nero d'Avola",
    "Petite Sirah",
    "Pinot Noir",
    "Port",
    "Portuguese Red",
    "Red Blend",
    "Rhône-style Red Blend",
    "Sangiovese",
    "Sangiovese Grosso",
    "Sauvignon",
    "Shiraz",
    "Syrah",
    "Tempranillo",
    "Tempranillo Blend",
    "Zinfandel"
  )
rose <- c("Rosé", "Rose")

wine_word_relation <- wine_mag_bing_sentiments %>%
  mutate(
    variety = case_when(
    variety %in% whites ~ "white",
    variety %in% reds ~ "red",
    variety %in% rose ~ "rose",
    TRUE ~ "other")) %>%
  group_by(word) %>%
  summarize(
    red = sum(variety == "red"),
    white = sum(variety == "white"),
    rose = sum(variety == "rose"),
    other = sum(variety == "other")) %>% 
  gather("variety","val",red,white,rose,other) %>% 
  filter(variety %in% c("red","white")) %>% 
  reshape2::acast(word~variety, value.var='val', fill = 0)


par(bg="black") 
wine_word_relation %>% 
  comparison.cloud(
    colors=c("#972726","#efe1a1"))




```
