---
title: "TidyX Episode 94 - Rmarkdown Guide - Parameterized Reports"
author: "TidyX"
output: html_document
params:
  demo_param1: "1000"
  demo_param2: 3000
---

## Parameters

As discussed ad nauseum, Rmarkdown facilitates reproducible reports that are able to
be rerun at the click of a button. 

But what if 
  - Some value change that needs to happen every time the report is run, but you want
   the same content (ie assumptions, model type, time period, etc)
  - Or your manager says "this is wonderful, but I want this report for 5 different company stock performances."

You _could_ copy and paste and have copies for every version or scenario.

Or, you could add parameters into your YAML header and call them rather than set
values in your document directly!


### YAML

The standard components of a yaml header are as follows:

 - title: the title of the document - string
 - author: who wrote the document - string
 - date: date report was created or rendered string
 - output: what is the rmarkdown output rendering to - function


For Parameterized Reports, a new component:

 - params: child content indicates the parameter values and how to set them
 
```yaml
params: 
  my_parameter_value: 10
  my_param_value: 20
```

Param values can also have child components, based on shiny input widgets for interactive
user inputs

```yaml
params: 
  my_parameter_value: 
   label: "My Param Input"
   value: 100
   type: numeric
   max: 1000
```

## Referencing params

The params values can be referenced using `params$param_name` in the code chunks just
as any other code.

```{r}
print(params$demo_param1)
print(params$demo_param2)
```

## Rendering with Params

- knit button
  - runs params as set. If selects "knit with param", attemps to select an input method based on type.
  
- Code via rmarkdown::render
  - argument "params" accepts a named list of params and values. if the param is not in the list, 
  the default value will be used:
  
```{r eval = FALSE}

rmarkdown::render(
  input = "This Report.Rmd", 
  output = "this report - 10.html"
  params = list(
    demo_param1 = "10"
  ))

rmarkdown::render(
  input = "This Report.Rmd", 
  output = "this report - 20.html"
  params = list(
    demo_param1 = "20"
  ))

```

## Resources

- Parameterized reports in R Markdown: The definitive Guide
  - https://bookdown.org/yihui/rmarkdown/parameterized-reports.html