---
title: "TidyX Episode 91 - Rmarkdown Guide - Chunk Options - Figures"
author: "TidyX"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

## Code Chunks

R code to be run is denoted as three (or more) ticks, followed by a curly brace, the letter R, a closing brace.
The area for the R code is in a new line, and continues until if finds the matching set of ticks (same number as initial set). 

```{r}

plot(cars)

lm(dist ~ speed, data = cars)

```

outputs of this code will be included by default unless options for the R chunks are set.

## Basic Code Chunk options

In the R chunk "header" many options can be set. Some basic and commonly used ones are:

- a *unique* chunk name, no spaces, suggested no underscores
- echo (boolean) - indicates whether code should be shared in the output
- include (boolean)- indicates whether the output or code should be included in the output (still run)
- eval (boolean) - indicates whether the code should be run
- warning/message (boolean) - indicates whether the warnings or messages output from the code should be displayed


```{r Regression-Cars, echo = FALSE, include = TRUE, eval = TRUE, warning = FALSE, message = TRUE}

lm(dist ~ speed, data = cars)

```


```{r setup, include=FALSE}

## set globally across Rmarkdown document from this point on
knitr::opts_chunk$set(
  echo = FALSE,
  include = TRUE,
  warning = FALSE, 
  message = FALSE
  )

```

## Figure Outputs

We can apply basic image output features to chunks where we expect to return an image/plot/etc.
These chunk options work across multiple output types. Prefixed with `fig.` in chunk options.

There are many options, these are the ones I find used most commonly

- fig.width
- fig.height
- fig.cap
- fig.align

```{r Plot-Cars, fig.width = 5, fig.cap = "Plot of the cars dataset", fig.align= 'center'}

plot(cars)

```

