---
title: "Episode 98 - Simulation & Resampling"
author: "Patrick Ward"
output: html_document
---

### Definitions

**Simulation:** A random generation of multiple samples from a defined data generating process.

**Resampling:** New data is *simulated* by resampling (with replacement) from the actual data set

**Mean:** A measure of central tendency. Represents the arithmetic average of the data.

**Standard Deviation:** A measure of dispersion/spread of the data relative to the mean.

**Standard Error:** Quantifies the accuracy of a sample measurement relative to a population. For example, the standard error of the mean for a sample of data represents the potential standard deviation of the mean in the population.  Always smaller than the standard deviation.


### Resampling

**Sample with replacement from a vector of numbers**

```{r}
d <- c(4, 2, 7, 9, 10, 1, 3, 6, 2, 7, 4)
sample(x = d, size = 5, replace = TRUE)
```


**Bootstrapping**

* Draw multiple random samples from observed data with replacement
* Draws must be independent and each observation must have an equal chance of being selected
* The bootstrap sample should be the same size as the observed data in order to use as much information from the sample as possible


*Let's look at Peyton Manning's Comp% during each of his NFL seasons*
- data courtesy of Pro-football-reference.com

```{r}
manning <- c(0.567, 0.621, 0.625, 0.627, 0.663, 0.67, 0.676, 0.673, 0.65, 0.654, 0.668, 0.688, 0.663, 0.686, 0.683, 0.662, 0.598)

mean(manning)                         # mean
sd(manning)                           # standard deviation
length(manning)                       # N samples
sd(manning) / sqrt(length(manning))   # standard error

hist(manning)

## Create a bootstrap sample of Manning's career production
set.seed(3)
n <- length(manning)
n.boot <- 1000

boot.manning <- rep(NA, n.boot)
head(boot.manning)

boot_set <- list()

for(i in 1:length(boot.manning)){
  # random sample of 1:n with replacement for the row indexes
  ind <- sample(1:n, replace = TRUE)
  
  # storing for future use (good practice)
  boot_set[[i]] <- ind
  
  # Use the row indexes to select the given values from the vector and calculate the mean
  boot.manning[i] <- mean(manning[ind])
}

head(boot.manning)
length(boot.manning)

## observe results
par(mfrow = c(1, 2))
hist(manning, main = "Peyton Manning\nObserved Comp%")
abline(v = mean(manning), col = "red", lwd = 3, lty = 2)
hist(boot.manning, main = "Peyton Manning\nBootstrap Comp%")
abline(v = mean(boot.manning), col = "red", lwd = 3, lty = 2)

### mean, sd, se, 95% CI
## original data
mean(manning)
sd(manning)
sd(manning) / sqrt(length(manning))
qnorm(p = c(0.025, 0.975), mean = mean(manning), sd =  sd(manning) / sqrt(length(manning)))

## bootstrapped data
# NOTE: because bootstrapping is the mean of means, the SD will be the standard error. Must convert back!
mean(boot.manning)
sd(boot.manning)                # SE
sd(boot.manning) * sqrt(n)      # Convert SE to SD
quantile(boot.manning, prob = c(0.025, 0.975))

```


*What if the sample has an extreme outlier? How can bootstrapping help?*

```{r}
# set the sample size (n) and the number of bootstrap samples (1000)
set.seed(8)
n <- 7
n.boot <- 1000

## create fake data with outlier
dat <- c(3, 1, 4, 2, 5, 3, 15)

mean(dat)
sd(dat)
hist(dat)

## bootstrap sample
boot.dat <- rep(NA, n.boot)
head(boot.dat)

# write for() loop for the resampling
for(i in 1:n.boot){
  # random sample of 1:n with replacement for the row indexes
  ind <- sample(1:n, replace = TRUE)
  
  # Use the row indexes to select the given values from the vector and calculate the mean
  boot.dat[i] <- mean(dat[ind])
}


# compare original data with bootstrap data
par(mfrow = c(1, 2))
hist(dat)
abline(v = mean(dat), col = "red", lwd = 3, lty = 2)
hist(boot.dat)
abline(v = mean(boot.dat), col = "red", lwd = 3, lty = 2)

### mean, sd, se, 95% CI
# original data
mean(dat)
sd(dat)
sd(dat) / sqrt(length(dat))
qnorm(p = c(0.025, 0.975), mean = mean(dat), sd =  sd(dat) / sqrt(length(dat)))

# bootstrapped data
# NOTE: because bootstrapping is the mean of means, the SD will be the standard error. Must convert back!
mean(boot.dat)
sd(boot.dat)             # SE
sd(boot.dat) * sqrt(n)   # Convert back to SD
quantile(boot.dat, prob = c(0.025, 0.975))
```


### Simulation

**Simulating a baseball hitter**

How many hits would we expect in 1 season of 500 ABs for a hitter with a 0.250 average?

```{r}
## parameters
season <- 1
AB <- 500
BA <- 0.250

rbinom(n = season, size = AB, prob = BA)         # count of hits
rbinom(n = season, size = AB, prob = BA) / AB    # batting average

```


**But that's only 1 simulation!!**

What if we re-ran that 1000 times? How much variability in performance would there be?


```{r}
# parameters
sims <- 1000
season <- 1
AB <- 500
BA <- 0.250

# vector to store results
results <- rep(NA, times = sims)

# for loop simulation
for(i in 1:sims){
  
  # re-run the rbinom() code 1000 times
  results[i] <- rbinom(n = 1, size = AB, prob = BA)
  
}

## plot results
par(mfrow = c(1,2))
hist(results, col = "light grey", main = "Distribution of the Number of Hits per Season")
abline(v = mean(results), col = "red", lwd = 2)
hist(results/AB, col = "light grey", main = "Distribution of the BA per Season")
abline(v = mean(results/AB), col = "red", lwd = 2)

## mean of simulated batting average
avg_ba <- mean(results/AB)
avg_ba

## standard deviation of batting average
sd_ba <- sqrt((avg_ba * (1 - avg_ba)) / length(results))
sd_ba

## 95% CI
avg_ba + qnorm(c(0.025, 0.975), mean = 0, sd = 1) * sd_ba

## Alternative approach to getting 95% CI
qnorm(c(0.025, 0.975), mean = avg_ba, sd = sd_ba) 

## 95% Quantile Interval
quantile(x = results/AB, probs = c(0.025, 0.975))

## Can do this simulation without without a for loop
ba_sim <- rbinom(n = sims, size = AB, prob = BA)
ba_sim

mean(ba_sim/AB)
```

