---
title: "Himalayan Climbing Expeditions"
author: "Patrick Ward"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(skimr)
library(plotly)
library(arm)
library(lme4)
library(merTools)

theme_set(theme_light())

tuesdata <- tidytuesdayR::tt_load('2020-09-22')

peaks <- tuesdata$peaks
exped <- tuesdata$expeditions
climbers <- tuesdata$members

```


### Climbers Data

```{r}

skim(climbers)

climbers %>%
  head()

```


### EDA

What type of role do people in each expedition have?

```{r}

climbers %>%
  count(expedition_role, sort = T) %>%
  top_n(5) %>%
  ggplot(aes(x = n, y = reorder(expedition_role, n))) +
  geom_col()

```


What sort of success has there been in climbing mountains?

```{r}

climbers %>%
  distinct(expedition_id, .keep_all = T) %>%
  count(peak_name, sort = T)

climbers %>%
  group_by(peak_name) %>%
  filter(n() >= 100) %>%
  count(peak_name, success) %>%
  group_by(peak_name) %>%
  mutate(Pct_Success = n / sum(n)) %>%
  filter(success == T) %>%
  arrange(desc(Pct_Success))

```


What is the count of deaths on the most common climbs?

```{r}

main_peaks <- climbers %>%
  distinct(expedition_id, .keep_all = T) %>%
  filter(n() >= 200) %>%
  count(peak_name, sort = T) %>%
  top_n(10) %>%
  pull(peak_name)

main_peak_exped <- climbers %>%
  filter(peak_name %in% main_peaks) %>%
  group_by(expedition_id) %>%
  mutate(Died = sum(died))

main_peak_exped %>%
  ggplot(aes(x = Died)) +
  geom_histogram()

exped_plt <- main_peak_exped %>%
  mutate(Decade = year -  year %% 10) %>%
  distinct(expedition_id, .keep_all = T) %>%
  group_by(Decade, peak_name) %>%
  summarize(
    Expeditions = n(),
    Deaths = sum(Died)
    ) %>%
  ggplot(aes(x = Decade, 
             y = Expeditions, 
             label = Deaths, 
             fill = peak_name,
             color = peak_name)) +
  geom_line(size = 1.1) +
  geom_point(size = 2.3,
             shape = 21,
             color = "black")

ggplotly(exped_plt)


```


Does the age of the climber influence the possibility of death on a climb?

```{r}

climber_deaths <- climbers %>%
  mutate(Role = fct_lump(expedition_role, 3),
         age_bin = cut_number(age, n = 4)) %>%
  filter(peak_name %in% main_peaks,
         !is.na(age))

table(climber_deaths$hired, climber_deaths$Role)

climber_deaths %>%
  ggplot(aes(x = age, fill = died)) +
  geom_density(alpha = 0.4)

climber_deaths %>%
  count(age_bin, died) %>%
  group_by(age_bin) %>%
  mutate(Pct = n / sum(n))

```


How does season influence death rate?

```{r}
climber_deaths %>%
  count(season, died) %>%
  group_by(season) %>%
  mutate(Pct = n / sum(n))
```


What is the death rate on some of the more population climbs?


```{r}

table(climber_deaths$peak_name, climber_deaths$died)

round(prop.table(table(climber_deaths$peak_name, climber_deaths$died), margin = 1) * 100, 2)

```


### Logistic Regression Mixed Model

Fit a logistic regression mixed model to the most common peaks with fixed effects for season, sex, oxygen used and climber role and allow the intercept to vary randomly for the peak.


```{r}

f_lmer <- glmer(
  died ~ season + sex + oxygen_used + Role + (1|peak_name),
  data = climber_deaths,
  family = "binomial"
  )

```

Model output

```{r}

display(f_lmer)

fixef(f_lmer)
ranef(f_lmer)
coef(f_lmer)

```

Coefficients are on a log scale. Transform coefficients back.

```{r}

exp(fixef(f_lmer))

```

Plot random effects


```{r}

lattice::dotplot(ranef(f_lmer, condVar = T))

```

## Make a prediction on the probability of death

Coefficients are interpreted on a relative scale. To understand outcomes on an absolute scale we can take the inverse logit of the entire equation.

New person:

* Climbing in Winter
* Sex = Female
* Oxygen Use = F
* Role = Leader
* Peak = Everest

```{r}

new_df <- data.frame(
        season = "Winter",
        sex = "F",
        oxygen_used = FALSE,
        Role = "Leader",
        peak_name = "Everest")

```


```{r}

inverse_logit <- function(x){
  prob <- 1 / (1 + exp(-x))
  prob
}

## get results for new person above
fixed <- fixef(f_lmer)[1] + fixef(f_lmer)[4] + fixef(f_lmer)[5]*0 + fixef(f_lmer)[6]*0 + fixef(f_lmer)[8]

# Use the inverse logit function to make a prediction on the fixed effects only
inverse_logit(fixed)

## Everest Effects
mtn <- ranef(f_lmer)$peak_name["Everest",]

## Adjust the fixed effects to account for Mt Everest
inverse_logit(fixed + mtn)

```


The same can be done a bit easier with the `predict()` function

```{r}
## Predict the outcome
predict(f_lmer, newdata = new_df,
  type = "link",
  re.form = NA,
        allow.new.levels = T,
        na.action = na.pass)

## run the result through the inverse logit function
inverse_logit(-4.246132 + 0.1623374)

```


Can use `predict()` to directly compute probabilities by setting the argument `type = "response"`

```{r}
predict(f_lmer, newdata = new_df,
  type = "response",
  re.form = ~(1|peak_name),
  allow.new.levels = T,
  na.action = na.pass)

```


What about prediction intervals?

```{r}

## prediction intervals
predictInterval(
  f_lmer, 
  level = 0.9, newdata = data.frame(
        season = "Winter",
        sex = "M",
        oxygen_used = FALSE,
        Role = "Climber",
        peak_name = "Everest"), 
    n.sims = 999, 
    type = "probability")


```


