---
title: "TidyX Episode 31 - Reactable"
author: "Patrick Ward & Ellis Hughes"
date: "October 9, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Packages & Data

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(reactable)
library(ggridges)
library(sparkline)
library(pagedtable) #thebioengineer/pagedtable

theme_set(theme_light())

```

```{r data}

tt <- tt_load("2020-10-06")

tt

ncaa <- tt$tournament

```


## EDA

View the data

```{r}

ncaa %>% 
  head() %>% 
  View()

```


How many conferences are in the data set?

```{r}

ncaa %>%
  count(conference, sort = T)

```

* Makes sense that the bigger conferences are represented in the tournament

* There are a lot of conferences and the top 6 are frequently referred to as the "Power 5" conferences.

Lump all conferences outside of the Power 5 (6 top conferences)

```{r}

ncaa_lumped <- ncaa %>%
  mutate(conf_lump = case_when(
    conference %in% c("Pacific-10","Pac-12") ~ "Pacific-10",
    conference %in% c(
      "Southeastern",
      "Atlantic Coast",
      "Big Ten",
      "Big East",
      "Big 12"
    ) ~ conference,
    TRUE ~ "Other"
  ))

```

Which of the Power-5 conferences have had the most tournament appearances?

```{r}

ncaa_lumped %>%
  filter(conf_lump != "Other") %>%
  count(conf_lump, sort = T) %>%
  mutate(Pct = n / sum(n)) %>%
  ggplot(aes(x = Pct, y = reorder(conf_lump, Pct))) +
  geom_col(fill = "light blue",
           alpha = 0.7,
           color = "black") +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "% of Appearances",
       y = "",
       title = "Percent of NCAA Women's Tournament Appearances Among Power-5 Schools")

```


Distribution of Regular Season Win/Loss Percentage Among Conferences

```{r}

ncaa_lumped %>%
  ggplot(aes(x = reg_percent/100, y = conf_lump, fill = conf_lump)) +
  geom_density_ridges(alpha = 0.7,
                      quantile_lines = T,
                      quantiles = 2) +
  scale_x_continuous(labels = scales::percent) +
  xlim(0,1)

```


Table

```{r table_prep}

ncaa_prepped <- ncaa_lumped %>%
  select(school, year, seed, reg_percent, tourney_w, tourney_l, tourney_finish) %>%
  group_by(school) %>% 
  summarize(
    
    school_summary = list(reg_percent),
    
    t_wins = sum(tourney_finish == "Champ"),
    
    tournament_wins = paste0(rep("&#x2606",sum(tourney_finish == "Champ")), collapse = ""),
    
    years = list(ungroup(.[.$school==unique(school),]))
    
  ) %>% 
  ungroup() %>% 
  arrange(desc(t_wins)) 

```

```{r table-making}

ncaa_prepped %>%
  
  select(school, tournament_wins, school_summary, years) %>% 
  
  reactable(
    
    columns = list(
      
      school = colDef(
        name = "School",
        filterable = TRUE,
        width = 100),
      
      tournament_wins = colDef(
        name = "Champs",
        html = TRUE,
        width = 80),
      
      school_summary = colDef(
        name = "Win % History",
        
        cell = function(value, index) {
          sparkline(
            ncaa_prepped$school_summary[[index]],
            width = "100px",
            height = "80px")
        },
        width = 120
      ),
      
      years = colDef(
        name = "Summary",
        
        cell = function(value, index) {
          
          pagedtable(
            ncaa_prepped$years[[index]] %>%
              select(-school),
            use_rownames = FALSE,
            height = "200px",
            pagerows = 5
          )
          
        }
      )
      
    ),
    
    bordered = TRUE,
    
    defaultPageSize = 5
  )
```

Win Probabilities

```{r}

ncaa_table <- ncaa %>%
  count(seed, tourney_finish) %>%
  mutate(
    tourney_finish = factor(
      tourney_finish,
      levels = c(
        "1st","2nd","RSF","RF","NSF","N2nd","Champ"
      )
    ),
  ) %>% 
  group_by(seed) %>%
  mutate(
    Pct = n / sum(n)
  )

ncaa_table %>%
  select(tourney_finish, Pct) %>%
  filter(tourney_finish != "OR",
         !is.na(seed)) %>%
  pivot_wider(names_from = "tourney_finish",
              values_from = "Pct") %>%
  select(seed, !!levels(ncaa_table$tourney_finish)) %>% 
  reactable(columns = list(
    '1st' = colDef(format = colFormat(percent = T, digits = 1), na = "-"),
    '2nd' = colDef(format = colFormat(percent = T, digits = 1), na = "-"),
    Champ = colDef(format = colFormat(percent = T, digits = 1), na = "-"),
    'N2nd' = colDef(format = colFormat(percent = T, digits = 1), na = "-"),
    NSF = colDef(format = colFormat(percent = T, digits = 1), na = "-"),
    RF = colDef(format = colFormat(percent = T, digits = 1), na = "-"),
    RSF = colDef(format = colFormat(percent = T, digits = 1), na = "-")
    ),
  defaultPageSize = 16)


```


