---
title: "TidyX Episode 107 - R Classes - Lists"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# List - what are they

- Object type that can contain literally any other object

- Objects are contained by element
  - elements can be identified by index number or name
  - multiple elements can be selected or removed at a time

- There are other, special types of lists in R (alist, pairlist) that we wont be getting into


```{r create-a-list}
## Create a list

list()

list(102540154,2,3,10,5,42)

list(
  c(1,2,3,4,5,6),
  
  data.frame(
    c1 = runif(10),
    c2 = 1:10
  ),
  
  list(
    1,2,3,4
  )
)

# Unnamed first element & named second element
first_list <- list(
  "TidyX list",
  "TidyX Numbers in a list" = c(1,3,5,7,9)
  )
first_list
```

Because lists are so flexible, there is a limited number of 

```{r list-properties}

# Basic Properties
length(first_list)
lengths(first_list)
names(first_list)

## lists with no names return null for names()
names(list(1,2,3))

```


Just like vectors, lists can be access through `[]` and `[[]]` calls.
 - `[]` returns the part of the list identified, and accepts vector inputs representing
 the indexes or names
 - `[[]]` returns the contents of the element, but can only be a length one number or character vector. if the element does not exist, it returns `NULL`
 - if there are named elements, `$` notation works

```{r}

# call elements
first_list[1]
first_list[2]

first_list[[2]]
first_list$`TidyX Numbers in a list`

# call the variables in the elements
first_list[2][[1]]

first_list[[2]]

# Call elements with tidyverse
first_list %>%
  .[1]

first_list %>%
  .[2]

first_list %>%
  .[2] %>%
  .[[1]]

# pluck() from purr to get the variables in a desired element
first_list %>%
  pluck(2)

```


# Iterating over lists

```{r}

set.seed(1234)
simple_list <- list(runif(10), runif(10), runif(10))

simple_list_mean <- list()

for(i in seq_along(simple_list)){
  simple_list_mean[[i]] <- mean(simple_list[[i]], na.rm = TRUE)
}


###### base R ############
# lapply()
lapply(X = simple_list, FUN = mean, na.rm = TRUE)

# sapply()
sapply(X = simple_list, FUN = mean, na.rm = TRUE, simplify = FALSE)
sapply(X = simple_list, FUN = mean, na.rm = TRUE, simplify = TRUE)

###### Tidyverse ############

simple_list %>% 
  map(mean, na.rm = TRUE)

simple_list %>% 
  map(~mean(.x, na.rm = TRUE))

simple_list %>% 
  setNames(c("run1","run2", "run3")) %>% 
  map(mean, na.rm = TRUE)

simple_list %>% 
  map_dbl(~mean(.x, na.rm = TRUE))

simple_list %>% 
  setNames(c("run1","run2", "run3")) %>% 
  map_dfr(~mean(.x, na.rm = TRUE))

simple_list %>% 
  map_chr(~mean(.x, na.rm = TRUE))

```




