---
title: "Naive Bayes pitchf/x Classification"
author: "TidyX"
date: "4/18/2021"
output: html_document
---

**Data courtest of {mlbgameday}:** https://github.com/keberwein/mlbgameday
**Resource for understanding pitchf/x data:** https://library.fangraphs.com/misc/pitch-fx/
**Resource on the features in the pitchf/x data:** Kagan, David. (2008). Fun with PitchFX Data. 


## Load Data

## References

**Data courtesy of {mlbgameday}:** https://github.com/keberwein/mlbgameday
**Resource for understanding pitchf/x data:** https://library.fangraphs.com/misc/pitch-fx/
**Resource on the features in the pitchf/x data:** Kagan, David. (2008). Fun with PitchFX Data. 

## Problem Statement

The FBL (fantasy baseball league) wants to predict which pitches are thrown. 
Traditionally it was labeled by a human sitting in the stands.
They want to productionalize it to give the classifications faster. 

They hired us...

## Load Data

```{r setup, include=FALSE}

bp <- here::here("TidyTuesday_Explained/057-MLB_pitch_classification_5")

knitr::opts_chunk$set(echo = TRUE,
                      root.dir = bp)

suppressPackageStartupMessages({
  suppressWarnings({
    library(tidyverse)
    library(xgboost)
    library(caret)
    library(doParallel)
  })
})


```

```{r load-data-and-clean}

## see Episode 53 - bit.ly/TidyX_Ep53 - for the background on this cleanup

# train
set1 <- readRDS(file.path(bp,"2016_04_21_to_2016_04_23_pitch.rds"))
set2 <- readRDS(file.path(bp,"2016_04_24_to_2016_04_26_pitch.rds"))
train <- bind_rows(set1, set2)

#test
test <- readRDS(file.path(bp,"2016_04_27_to_2016_04_30_pitch.rds"))

names(train)

#### Cleaning data based on EDA (below) ----------------------------------------------
train_cleaned <- train %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

test_cleaned <- test %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO", "SC"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

## Check for any NA's

train_cleaned %>%
  summarize(across(.cols = everything(),
                  .fns = ~sum(is.na(.x)))) %>%
  t()

test_cleaned %>%
  summarize(across(.cols = everything(),
                  .fns = ~sum(is.na(.x)))) %>%
  t()


train_cleaned$pitch_type <- as.factor(train_cleaned$pitch_type)
test_cleaned$pitch_type <- as.factor(test_cleaned$pitch_type)
```


## Naieve Bayes Classifier

```{r}
## convert pitch type to numeric
train_cleaned$pitch_type_num <- as.numeric(train_cleaned$pitch_type)-1
test_cleaned$pitch_type_num <- as.numeric(test_cleaned$pitch_type)-1

table(train_cleaned$pitch_type, train_cleaned$pitch_type_num)

library(e1071)

fit_nb <- naiveBayes(x = train_cleaned %>% select(-pitch_type, -pitch_type_num),
                     y = train_cleaned %>% select(pitch_type_num))

preds <- predict(fit_nb, newdata = test_cleaned, type = "raw") %>% 
  data.frame() %>% 
  setNames(levels(test_cleaned$pitch_type)) %>% 
  mutate(
    across(everything(), round, 3)
  )

# Identify the highest probability
preds$pred_pitch <- factor(
  levels(test_cleaned$pitch_type)[max.col(preds, ties.method = "first")],
  levels = levels(test_cleaned$pitch_type))


# Put the original pitch types into the predicted data frame
preds$original_pitch_type <- test_cleaned$pitch_type

preds %>% head()

## confusion matrix
table(preds$original_pitch_type, preds$pred_pitch,
      dnn = c("Observed", "Predicted"))


sum(diag(table(preds$original_pitch_type, preds$pred_pitch))) / nrow(preds)

# within pitch type accuracy

preds %>%
  count(pred_pitch, original_pitch_type) %>%
  group_by(original_pitch_type) %>%
  summarize(N = sum(n),
            matches = sum(ifelse(pred_pitch == original_pitch_type, n, 0)),
    within_acc = matches / N)

```

