---
title: "Imputing NAs"
author: "Patrick Ward"
date: "12/4/2020"
output: html_document
---

Previous blog article written about simple approaches to dealing with NA, NaN, and Inf, in a data frame: http://optimumsportsperformance.com/blog/r-tips-tricks-dealing-with-missing-values/


## Load Packages and Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Packages
library(tidyverse)
library(janitor)
library(patchwork)
library(rvest)

theme_set(theme_light())

## data

url <- read_html("https://www.pro-football-reference.com/draft/2013-combine.htm")

combine <- url %>%
  html_table(fill = T) %>%
  purrr::pluck(1) %>%
  filter(Player != "Player") %>%
  clean_names() %>%
  mutate(
    across(c(6:12), .fns = ~readr::parse_number(.x)),
    hgt_ft = as.numeric(gsub("-.*", "", ht))*12,
    hgt_in = as.numeric(gsub("*.-", "", ht)),
    ht = hgt_ft + hgt_in,
    pos_group = case_when(pos %in% c("OT", "OG", "C") ~ "OL",
                          pos %in% c("ILB", "OLB") ~ "LB",
                          pos %in% c("S", "CB") ~ "DB",
                          pos == "EDGE" ~ "DE",
                          pos == "DT" ~ "DL",
                          TRUE ~ as.character(pos))
  ) %>%
  select(-hgt_ft, -hgt_in) %>%
  separate(drafted_tm_rnd_yr, into = c("Tm", "Rd", "Pick", "Yr"), sep = "/") %>%
  mutate(Round = substring(Rd, 1, 2),
         Round = ifelse(is.na(Round), "FA", as.character(Round)),
         Year = ifelse(is.na(Yr), 2013, Yr)) %>%
  select(-Rd, -Pick, -Yr) %>%
  filter(!pos %in% c("K", "LS", "P"))

combine %>% head()
combine %>% glimpse()

```


## How many NAs in each columns?

```{r}
combine %>%
  summarize(across(everything(), ~sum(is.na(.x))))

```


Are the missing performance values specific to different position groups?

```{r}
combine %>%
  group_by(pos_group) %>%
  summarize(
    across(everything(), ~sum(is.na(.x))),
     n = n()
    ) %>% 
  relocate(n, .after = pos_group) %>% 
  mutate(
    across(player:Year, ~ .x/n)
  )

```


## Mean and Median Imputation

* Impute values for the 3cone test

```{r}

combine_imp <- combine %>%
  mutate(
    needs_imputation = ifelse(is.na(x3cone), 1, 0),
    mean_imp = ifelse(is.na(x3cone), mean(x3cone, na.rm = T), x3cone),
    median_imp = ifelse(is.na(x3cone), median(x3cone, na.rm = T), x3cone)
  )

```


* What if we do it within position group instead?

```{r}

combine_imp <- combine_imp %>%
  group_by(pos_group) %>%
  mutate(
    mean_imp_pos = ifelse(is.na(x3cone), mean(x3cone, na.rm = T), x3cone),
    median_imp_pos = ifelse(is.na(x3cone), median(x3cone, na.rm = T), x3cone)
  ) %>% 
  ungroup()

```


```{r}

combine_imp %>% head()

```


## Impute using linear regression

* build a regression model to predict 3cone from wt and x40yd

```{r}
## 3cone ~ Wt
combine_imp %>%
  ggplot(aes(x = wt, y = x3cone, color = pos_group)) +
  geom_point(size = 4,
             alpha = 0.8) +
  labs(title = "3cone ~ Wt",
     x = "Wt",
     y = "3cone")

## 3cone ~ 40yd Sprint
combine_imp %>%
  ggplot(aes(x = x40yd, y = x3cone, color = pos_group)) +
  geom_point(size = 4,
             alpha = 0.8) +
  labs(title = "3cone ~ 40yd Sprint",
     x = "40yd Sprint",
     y = "3cone")

# model
fit_three_cone <- lm( x3cone ~ wt + x40yd , data = combine_imp)
summary(fit_three_cone)

# add predictions into the data
combine_imp <- combine_imp %>%
  mutate(lm_three_cone = ifelse(is.na(x3cone), 
                                predict(fit_three_cone, newdata = .),
                                x3cone))

combine_imp %>% head(20)

# histogram of imputed values
hist(combine_imp$lm_three_cone, col = "light grey", main = "Imputed 3cone (Linear Model)", xlab = "3cone Impuated")

# Densities by position groups
combine_imp %>%
  mutate(side = case_when(pos_group %in% c("DB", "DE", "DL", "LB") ~ "Defense",
                          TRUE ~ "Offense")) %>%
  ggplot(aes(x = lm_three_cone, fill = pos_group)) +
  geom_density(alpha = 0.4) +
  scale_color_binned() +
  facet_wrap(~side)

```

## MICE package

```{r mice}

library(mice)

md.pattern(combine)

combine_mice <- combine %>% 
  mice(m = 5, seed = 12345) %>% 
  {
    
    three_cone_imp <- .$imp$x3cone %>% 
      pull(5)
    
    dat <- .$data %>% 
      mutate(
        needs_imputation = ifelse(is.na(x3cone), 1, 0)
        )
    
    dat[is.na(dat$x3cone), "x3cone"] <- three_cone_imp
    
    dat
  }
                     
```

* looking at the relationship between different kinds of imputation

```{r}


combine_imp %>%
  dplyr::mutate(mice_three_cone = combine_mice$x3cone) %>%
  select(lm_three_cone,
         mean_imp_pos,
         median_imp_pos,
         mice_three_cone,
         x3cone,
         needs_imputation) %>%
  pivot_longer(
    cols = c(lm_three_cone, mean_imp_pos, median_imp_pos, mice_three_cone, x3cone),
    names_to = "imputation",
    values_to = "time"
  ) %>%
  filter(
    imputation %in% c("lm_three_cone", "mean_imp_pos","median_imp_pos", "mice_three_cone") &
      needs_imputation == 1 |
      imputation %in% c("x3cone") & needs_imputation == 0
  ) %>%
  mutate(
    imputation = factor(
      imputation,
      levels = c("mean_imp_pos","median_imp_pos", "lm_three_cone", "mice_three_cone", "x3cone")
    )
  ) %>% 
  ggplot(aes(x = time,
             fill = imputation)) +
  geom_density() +
  facet_grid( ~ imputation) +
  labs(x = "3 cone",
       title = "Imputed & Non-Imputed 3-Cone")

```


