---
title: "pitchf/x class imbalance"
author: "TidyX"
date: "4/21/2021"
output: html_document
---

**Data courtest of {mlbgameday}:** https://github.com/keberwein/mlbgameday
**Resource for understanding pitchf/x data:** https://library.fangraphs.com/misc/pitch-fx/
**Resource on the features in the pitchf/x data:** Kagan, David. (2008). Fun with PitchFX Data. 


## Load Data

```{r setup, include=FALSE}

bp <- here::here("TidyTuesday_Explained/059-MLB_pitch_classification_7")

knitr::opts_chunk$set(echo = TRUE,
                      root.dir = bp)

suppressPackageStartupMessages({
  suppressWarnings({
    library(tidyverse)
    library(randomForest)
    library(caret)
  })
})

theme_set(theme_light())

set1 <- readRDS(file.path(bp,"2016_04_21_to_2016_04_23_pitch.rds"))
set2 <- readRDS(file.path(bp,"2016_04_24_to_2016_04_26_pitch.rds"))
train <- bind_rows(set1, set2)

#test
test <- readRDS(file.path(bp,"2016_04_27_to_2016_04_30_pitch.rds"))

names(train)


#### Cleaning data based on EDA (below) ----------------------------------------------
train_cleaned <- train %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

test_cleaned <- test %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO", "SC"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

## Check for any NA's

train_cleaned %>%
  summarize(across(.cols = everything(),
                  .fns = ~sum(is.na(.x)))) %>%
  t()

test_cleaned %>%
  summarize(across(.cols = everything(),
                  .fns = ~sum(is.na(.x)))) %>%
  t()


train_cleaned$pitch_type <- as.factor(train_cleaned$pitch_type)
test_cleaned$pitch_type <- as.factor(test_cleaned$pitch_type)
```


How much imbalance?

```{r}
train_cleaned %>%
  count(pitch_type) %>%
  mutate(pct = n/ sum(n)) %>%
  ggplot(aes(x = n, y = reorder(pitch_type, n))) +
  geom_col(fill = "lightblue") +
  geom_label(aes(label = scales::percent(pct)))
```



Run the original random forest [Episode 55](bit.ly/TidyX_Ep55)


```{r}

set.seed(567)
fit_rf <- randomForest(pitch_type ~ ., data = train_cleaned, mtry = 6)

test_rf_pred <- bind_cols(test_cleaned, rf_pred = predict(fit_rf, newdata = test_cleaned, type = "class"))
test_rf_pred %>% head()

test_rf_pred %>%
  janitor::tabyl(rf_pred, pitch_type)

test_rf_pred %>%
  count(rf_pred, pitch_type) %>%
  group_by(pitch_type) %>%
  summarize(N = sum(n),
            matches = sum(ifelse(pitch_type == rf_pred, n, 0)),
            within_acc = matches / N)

## overall accuracy
sum(diag(table(test_rf_pred$rf_pred, test_rf_pred$pitch_type))) / sum(table(test_rf_pred$rf_pred, test_rf_pred$pitch_type))
```



Create down and up sampled data sets for modeling

```{r}

sort(table(train_cleaned$pitch_type))

set.seed(5873)
?downSample
train_down <- downSample(x = train_cleaned[, -1],
                     y = train_cleaned$pitch_type)

sort(table(train_down$Class))

set.seed(5873)
train_up <- upSample(x = train_cleaned[, -1],
                     y = train_cleaned$pitch_type)

sort(table(train_up$Class))

```


## Build new models on down and up sampled data

down sampled model

```{r}

set.seed(3232)
fit_rf_down <- randomForest(Class ~ ., data = train_down, mtry = 6)

test_rf_pred <- bind_cols(test_rf_pred, rf_pred_down = predict(fit_rf_down, newdata = test_rf_pred, type = "class"))
test_rf_pred %>% head()

test_rf_pred %>%
  janitor::tabyl(rf_pred_down, pitch_type)

test_rf_pred %>%
  count(rf_pred_down, pitch_type) %>%
  group_by(pitch_type) %>%
  summarize(N = sum(n),
            matches = sum(ifelse(pitch_type == rf_pred_down, n, 0)),
            within_acc = matches / N)

## overall accuracy
sum(diag(table(test_rf_pred$rf_pred_down, test_rf_pred$pitch_type))) / sum(table(test_rf_pred$rf_pred_down, test_rf_pred$pitch_type))

```


up sampled model

```{r}

set.seed(3232)
fit_rf_up <- randomForest(Class ~ ., data = train_up, mtry = 6)

test_rf_pred <- bind_cols(test_rf_pred, rf_pred_up = predict(fit_rf_up, newdata = test_rf_pred, type = "class"))
test_rf_pred %>% head()

test_rf_pred %>%
  janitor::tabyl(rf_pred_up, pitch_type)

test_rf_pred %>%
  count(rf_pred_up, pitch_type) %>%
  group_by(pitch_type) %>%
  summarize(N = sum(n),
            matches = sum(ifelse(pitch_type == rf_pred_up, n, 0)),
            within_acc = matches / N)

## overall accuracy
sum(diag(table(test_rf_pred$rf_pred_up, test_rf_pred$pitch_type))) / sum(table(test_rf_pred$rf_pred_up, test_rf_pred$pitch_type))
```


## Model Comparison

Log-Loss, Brier Score, ROC

```{r}
library(MLmetrics) # for log-loss
library(measures) # for multi-class brier score
library(pROC) # for ROC

 # add probabilities for each model to the data set
test_rf_pred <- bind_cols(test_rf_pred, rf_pred_prob = predict(fit_rf, newdata = test_cleaned, type = "prob"))
test_rf_pred <- bind_cols(test_rf_pred, rf_pred_down_prob = predict(fit_rf_down, newdata = test_rf_pred, type = "prob"))
test_rf_pred <- bind_cols(test_rf_pred, rf_pred_up_prob = predict(fit_rf_up, newdata = test_rf_pred, type = "prob"))

test_rf_pred %>% head()

# calculate log loss
MultiLogLoss(y_true = test_rf_pred$pitch_type, y_pred = test_rf_pred$rf_pred_prob)
MultiLogLoss(y_true = test_rf_pred$pitch_type, y_pred = test_rf_pred$rf_pred_down_prob)
MultiLogLoss(y_true = test_rf_pred$pitch_type, y_pred = test_rf_pred$rf_pred_up_prob)

# dumb log loss for balanced data
n_classes <- length(unique(test_rf_pred$pitch_type))

-log(1/n_classes)

## Brier score
multiclass.Brier(truth = test_rf_pred$pitch_type, probabilities = test_rf_pred$rf_pred_prob)
multiclass.Brier(truth = test_rf_pred$pitch_type, probabilities = test_rf_pred$rf_pred_down_prob)
multiclass.Brier(truth = test_rf_pred$pitch_type, probabilities = test_rf_pred$rf_pred_up_prob)

## ROC

suppressMessages({
  multiclass.roc(response = test_rf_pred$pitch_type,
    predictor = as.numeric(test_rf_pred$rf_pred))$auc
  })

suppressMessages({
  multiclass.roc(response = test_rf_pred$pitch_type,
    predictor = as.numeric(test_rf_pred$rf_pred_down))$auc
  })

suppressMessages({
  multiclass.roc(response = test_rf_pred$pitch_type,
    predictor = as.numeric(test_rf_pred$rf_pred_up))$auc
  })

```
