---
title: "TidyX Episode 106 - R Classes - Factors"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Factors - what are they

- `factor()`

- Provides a way to provide different categories/enumerations of things without specifying a 
  priority preference. 
    - IE. in `mtcars` car manufacturers may be a included in a regression, but 
    order in which they are included in the model shouldn't impact the results. 
    Or Ferrari is not greater than Mazda from a regression standpoint.
    
- Option for providing ordinality when it is important to sort categories
    - IE First, Second, Third, are categories, but there is an order to them

- Until R4.0.0, stringsAsFactors was default `TRUE`when loading data with 
    base R functions (read.table) or creating a data.frame

```{r creation-of-factors, eval=FALSE}

## Factors can be build from other vectors
factor(c("Hello","world","Hello","TidyX"))
factor(c(TRUE,FALSE,TRUE))
factor(c(1,2,3))

## You can specify the levels that exist in the factor vector
factor(c("B","A","C","D","A","B"), levels = LETTERS)

## Check factor levels using `levels()`
ex_f <- factor(c("Hello","world","Hello","TidyX"))
levels(ex_f)

## You can specify whether there is an ordinality to the levels
## order is Increasing when ordinal
factor(c("B","A","C","D","A","B"), levels = LETTERS, ordered = TRUE)
factor(c("B","A","C","D","A","B"), levels = rev(LETTERS), ordered = TRUE)

## Factor structure
dput(factor(c("Hello","world","Hello","TidyX")))

```

Factors can be a memory-efficient way to store your vectors where repeated values are common:

```{r}

large_char_vector <- sample(
  c("TEST","THIS","IS","A","LARGE","NUMBER","OF","WORDS"),
  100000,
  replace = TRUE
  )

large_char_vector_f <- as.factor(large_char_vector)

print(object.size(large_char_vector), units = "MB")
print(object.size(large_char_vector_f), units = "MB")

```

Levels of factors are the important piece to them, and they can be edited

```{r}

vec <- factor(c("Hello","world","Hello","TidyX"))
vec

factor(c("Hello","world","Hello","TidyX"), levels = c("world","Hello"))

levels(vec) <- c("Amazing","Viewers","Are")
vec

## fails due to not having at least the same number of levels
levels(vec) <- c("Check","the TidyX")

```

# Factor Conversion

Conversion into factors is relatively simply, but conversion back is not as simple
for all types (numeric)

```{r}

char_f <- factor(c("Hello","world","Hello","TidyX"))
bool_f <- factor(c(TRUE,FALSE,TRUE))
num_f  <- factor(c(3,30,2,1,10,4,1000))

## character works as expected b/c it has a method for conversion
as.character(char_f)

## boolean works as expected b/c it has a method for conversion
as.logical(bool_f)

## as.numeric extract the numeric representations of the factors
as.numeric(num_f)

## conversion to character then numeric is how you can extract back to numeric
as.numeric(as.character(num_f))

```

# Factors Uses

Regression

```{r}

library(tidyverse)

mtcars_f <- mtcars %>% 
  mutate(
    cyl_f_ = factor(cyl),
    cyl_f2_ = factor(cyl, levels = c("8","6","4"))
  )

lm_num <- lm(mpg ~ cyl, data = mtcars_f)
lm_f <- lm(mpg ~ cyl_f_, data = mtcars_f)
lm_f2 <- lm(mpg ~ cyl_f2_, data = mtcars_f)
  
lm_num
lm_f
lm_f2

summary(lm_num)
summary(lm_f)
summary(lm_f2)



tibble(
  lm_num_pred_num = predict(lm_num, newdata = data.frame(cyl = 3.6)),
  lm_num_pred_int = predict(lm_num, newdata = data.frame(cyl = 4)),
  lm_f_pred = predict(lm_f, newdata = data.frame(cyl_f_ = "4")),
  lm_f2_pred = predict(lm_f2, newdata = data.frame(cyl_f2_ = "4"))
)

```

ggplot bar order

```{r}

ggplot(mtcars_f) +
  geom_boxplot(
    aes(x = cyl, y = mpg)
  )

ggplot(mtcars_f) +
  geom_boxplot(
    aes(x = cyl_f_, y = mpg)
  )

ggplot(mtcars_f) +
  geom_boxplot(
    aes(x = cyl_f2_, y = mpg)
  )

````


