---
title: "TidyX Episode 9: Tables"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup}
library(tidyverse)
library(rvest)
```

## Data Source

Data that we are looking at this week comes directly from the publically available data from the NFl. In this case, we will be looking at the kickers of the 2019 season. 

```{r load-data}
#### Data Loading & Clean Up ---------------------------------------------------------------

## get data table
kickers <- read_html("http://www.nfl.com/stats/categorystats?archive=false&conference=null&statisticCategory=FIELD_GOALS&season=2019&seasonType=REG&experience=&tabSeq=0&qualified=false&Submit=Go") %>%
  html_nodes("table") %>%
  html_table() %>%
  purrr::pluck(1)

kickers %>% head()
names(kickers)
```

If you take a close look at the URL, you can see there are pieces that look like they are arguments, similar to R. We can use those to traverse statistics across 
multiple years to get more information. In this episode, we will limit the analysis to the 2019 season.

## Data Cleaning

After the reading a table from a website, it can have formatted elements that make it difficult to analyze initially. For this example, it included 2 levels of headers, where the upper level headers that cross multiple margins.

To handle this, we need to clean up and make column names that are all unique.

```{r clean-data}

kickers_rename <- kickers

## Get columns for analysis
row1 <- kickers_rename[1,]
colnames(kickers_rename) <- row1
kickers_rename <- kickers_rename[-1,]

kickers_clean <- kickers_rename[,c(2,3,5,6,10,12,14,16,18)]
colnames(kickers_clean) <- c("Player", 
                       "Team", 
                       "Made", 
                       "Att", 
                       "X1_to_19",
                       "X20_to_29",
                       "X30_to_39",
                       "X40_to_49",
                       "X50_plus")

```

```{r clean-data2, eval = FALSE}

kickers_rename <- kickers

## Set column names for analysis
colnames(kickers_rename) <- trimws(paste( kickers_rename[1,], colnames(kickers_rename)))
kickers_rename <- kickers_rename[-1,]

kickers_clean <- kickers_rename %>% 
  select(
    Player, 
    Team, 
    Made = `FGM FG Overall`,
    Att = `FG Att FG Overall`,
    contains("A-M")
  )

```


```{r process-data}
kickers_statline <- kickers_clean %>%
  separate(X1_to_19, into = c("X1_to_19_Att", "X1_to_19_Made"), sep = "-") %>%
  separate(X20_to_29, into = c("X20_to_29_Att", "X20_to_29_Made"), sep = "-") %>%
  separate(X30_to_39, into = c("X30_to_39_Att", "X30_to_39_Made"), sep = "-") %>%
  separate(X40_to_49, into = c("X40_to_49_Att", "X40_to_49_Made"), sep = "-") %>%
  separate(X50_plus, into = c("X50_plus_Att", "X50_plus_Made"), sep = "-")

## Turn characters into numeric
numeric_cols <- c(3:14)
kickers_statline[, numeric_cols] <-
  apply(kickers_statline[, numeric_cols], 2, function(x)
    as.numeric(as.character(x)))

```

```{r process-data2, eval = FALSE}
## Separate strings

kickers_statline <- kickers_clean %>%
  gather("Distance","statline",contains("A-M")) %>% 
  separate(statline, into = c("A","M"), sep = "-") %>% 
  mutate(
    Att = as.numeric(Att),
    Made = as.numeric(Made),
    A = as.numeric(A),
    M = as.numeric(M),
    Distance = gsub("A-M\\s+","",Distance)
  )

glimpse(kickers_statline)

```

```{r adjusting}

## Distance Adjustment ------------------------------------------------------------

X1_to_19 <- with(kickers_statline, sum(X1_to_19_Att, na.rm = T) / sum(Att, na.rm = T))
X20_to_29 <- with(kickers_statline, sum(X20_to_29_Att, na.rm = T) / sum(Att, na.rm = T))
X30_to_39 <- with(kickers_statline, sum(X30_to_39_Att, na.rm = T) / sum(Att, na.rm = T))
X40_to_49 <- with(kickers_statline, sum(X40_to_49_Att, na.rm = T) / sum(Att, na.rm = T))
X50_plus <- with(kickers_statline, sum(X50_plus_Att, na.rm = T) / sum(Att, na.rm = T))

X1_to_19
X20_to_29
X30_to_39
X40_to_49
X50_plus

## Apply Adjsutment to each player ---------------------------------------
kickers_adj <- kickers_statline

kickers_adj$X1_to_19_adj <- ifelse(kickers_adj$X1_to_19_Att > 0, 
                                   kickers_adj$X1_to_19_Made / kickers_adj$X1_to_19_Att * X1_to_19,
                                   0)
kickers_adj$X20_to_29_adj <- ifelse(kickers_adj$X20_to_29_Att > 0, 
                                    kickers_adj$X20_to_29_Made / kickers_adj$X20_to_29_Att * X20_to_29, 
                                    0)
kickers_adj$X30_to_39_adj <- ifelse(kickers_adj$X30_to_39_Att > 0, kickers_adj$X30_to_39_Made / kickers_adj$X30_to_39_Att * X30_to_39, 0)
kickers_adj$X40_to_49_adj <- ifelse(kickers_adj$X40_to_49_Att > 0, kickers_adj$X40_to_49_Made / kickers_adj$X40_to_49_Att * X40_to_49, 0)
kickers_adj$X50_plus_adj <- ifelse(kickers_adj$X50_plus_Att > 0, kickers_adj$X50_plus_Made / kickers_adj$X50_plus_Att * X50_plus, 0)

Kickers_Att <- kickers_adj

Kickers_Att$Adj_FG_Pct <- with(Kickers_Att, X1_to_19_adj + X20_to_29_adj + X30_to_39_adj + X40_to_49_adj + X50_plus_adj)
Kickers_Att$UnAdj_FG_Pct <- with(Kickers_Att, Made / Att)
head(kickers)

```

```{r adjusting2, eval = FALSE}
## Distance Adjustment ------------------------------------------------------------

Kickers_Att <- kickers_statline %>% 
  group_by(Distance) %>% 
  summarize(
    Att_Percent = sum(A, na.rm = TRUE )/sum(Att, na.rm = TRUE)
  )

Kickers_Att

## Apply Adjsutment to each player ---------------------------------------

Kickers_Adj <- kickers_statline %>% 
  left_join(Kickers_Att, by = "Distance") %>% 
  mutate(
    M_adj = ifelse( A > 0, M/A * Att_Percent, 0 )
  ) %>% 
  group_by(Player, Team, Made, Att) %>% 
  summarize(
    Adj_FG_Pct = sum(M_adj),
    UnAdj_FG_Pct = sum(M) / sum(A)
  ) %>% 
  ungroup()

```

## Different approaches to creating tables

```{r gt}

library(gt)

## gt package
Kickers_Att %>%
  mutate(Adj_FG_Pct = round(Adj_FG_Pct * 100, 1),
         UnAdj_FG_Pct = round(UnAdj_FG_Pct * 100, 1)) %>%
  arrange(desc(Adj_FG_Pct)) %>%
  select(Player, Team, 'Overall Att' = Att, 'Adj FG%' = Adj_FG_Pct, 'Raw FG%' = UnAdj_FG_Pct) %>% 
  gt()

```


```{r condformat}
## condformat package

library(condformat)

Kickers_Att %>%
  mutate(Adj_FG_Pct = round(Adj_FG_Pct * 100, 1),
         UnAdj_FG_Pct = round(UnAdj_FG_Pct * 100, 1)) %>%
  arrange(desc(Adj_FG_Pct))  %>%
    select(Player, Team, 'Overall Att' = Att, 'Adj FG%' = Adj_FG_Pct, 'Raw FG%' = UnAdj_FG_Pct) %>% 
  
  condformat() %>%
  rule_fill_gradient(`Adj FG%`, 
                     low = "blue", 
                     high = "green")
```

```{r colortable}
## colortable package

# remotes::install_github("thebioengeer/colortable)
library(colortable)

Kickers_Att  %>% 
  tibble() %>%
  mutate(Adj_FG_Pct = round(Adj_FG_Pct * 100, 1),
         UnAdj_FG_Pct = round(UnAdj_FG_Pct * 100, 1)) %>%
  arrange(desc(Adj_FG_Pct))  %>%
  mutate(Adj_FG_Pct = color_vctr(Adj_FG_Pct, text_color = color_scale(colorRamp(c(
    "blue", "green"
  ))))) %>%
  select(Player, Team, 'Overall Att' = Att, 'Adj FG%' = Adj_FG_Pct, 'Raw FG%' = UnAdj_FG_Pct) %>% 
  knitr::kable()

```
