---
title: "TidyX 78 - tidyverse model framework"
author: "Patrick Ward"
output: html_document
---

## Load packages & data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(patchwork)
library(datasets)

theme_set(theme_light())

## function set up
plot_continuous <- function(col){
  plt <- df %>%
    ggplot(aes(x = {{col}}, y = Ozone)) +
    geom_point() +
    geom_smooth(method = "lm")
  plt
}

plot_categorical <- function(col){
  plt <- df %>%
    ggplot(aes(x = {{col}}, Ozone, y = Ozone)) +
    geom_boxplot()
  plt
}

```


```{r data}

data("airquality")

?airquality

df <- airquality %>%
  mutate(Month.f = as.factor(Month))

df %>%
  head()

df %>%
  summarize(across(everything(),
            ~sum(is.na(.x))))

```


## EDA

```{r}

df %>%
  nrow()

(plot_continuous(Solar.R) + plot_continuous(Wind)) /
  (plot_continuous(Temp) + plot_categorical(Month.f))
```


## Building some tidymodels to predict Ozone

The main packages in the `tidymodels` suite that we will use:


`rsample`: train/test or cross validation splits
`recipes`: pre-processing of data
`parsnip`: specifying the model we want to use
`workflows` : for putting the model together and constructing the workflow
`yardstick`: model evaluation metrics
`broom` : for model outputs in a clean/tidy data frame

**Data splitting with `rsample`**


```{r test-train-split}

df_split <- initial_split(df,prop = 3/4, strata = Month.f)
df_split

df_train <- training(df_split)
df_test <- testing(df_split)

```


**Create a model recipe**

* Setting the recipe is just specifying the model.
* Key note here is that there are a number of helper functions (e.g., `step_normalize()`, `step_impute_median()`, `step_dummy()`) that can be useful for handling data preparation. You would do all pre-processing in this step.

```{r}

aq_recipe <- 
  recipe(
    Ozone ~ Solar.R + Wind + Temp + Month.f ,
    data = df_train
  ) %>%
  step_impute_median(Ozone, Solar.R) %>% 
  step_normalize(Solar.R, Wind, Temp)

aq_recipe

```


**Initialize the model**

```{r}

lm_model <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

lm_model
```


**Set up the workflow**

This step is initiated with `workflow()` function

```{r}
# set the workflow
aq_workflow <-
  workflow() %>%
  add_model(lm_model) %>%
  add_recipe(aq_recipe) 

aq_workflow
```


**Run the model on our training data set**

```{r}

model_fit <- aq_workflow %>% 
  fit(data = df_train)

```


**Review Model

```{r}

model_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

```

**Run Model on Test Set**

```{r}

ozone_pred <- predict(
  model_fit %>% extract_fit_parsnip(),
  model_fit %>% extract_recipe() %>% bake(df_test)
  )

fit_df_test <- bind_cols(df_test, ozone_pred)

```


```{r}

## plot
fit_df_test %>%
  ggplot(aes(x = .pred, y = Ozone)) +
  geom_point(size = 6,
             aes(color = Month.f),
             alpha = 0.5) +
  geom_abline(intercept = 0,
              slope = 1,
              size = 1.5,
              linetype = "dashed")

```

