---
title: "pitchf/x analysis"
author: "Patrick Ward"
date: "3/18/2021"
output: html_document
---

## References

**Data courtesy of {mlbgameday}:** https://github.com/keberwein/mlbgameday
**Resource for understanding pitchf/x data:** https://library.fangraphs.com/misc/pitch-fx/
**Resource on the features in the pitchf/x data:** Kagan, David. (2008). Fun with PitchFX Data. 

## Problem Statement

The FBL (fantasy baseball league) wants to predict which pitches are thrown. 
Traditionally it was labeled by a human sitting in the stands.
They want to productionalize it to give the classifications faster. 

They hired us...

## Load Data

```{r setup, include=FALSE}

bp <- here::here("TidyTuesday_Explained/053-MLB_pitch_classification_introduction")

knitr::opts_chunk$set(echo = TRUE,
                      root.dir = bp)

suppressPackageStartupMessages({
  suppressWarnings({
    library(tidyverse)
    library(corrplot)
    library(plotly)
  })
})

theme_set(theme_light())

# train
set1 <- readRDS(file.path(bp,"2016_04_21_to_2016_04_23_pitch.rds"))
set2 <- readRDS(file.path(bp,"2016_04_24_to_2016_04_26_pitch.rds"))
train <- bind_rows(set1, set2)

#test
test <- readRDS(file.path(bp,"2016_04_27_to_2016_04_30_pitch.rds"))

names(train)


#### Cleaning data based on EDA (below) ----------------------------------------------
train_cleaned <- train %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

test_cleaned <- test %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO", "SC"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0)

## Check for any NA's

train_cleaned %>%
  summarize(across(.cols = everything(),
                  .fns = ~sum(is.na(.x)))) %>%
  t()

test_cleaned %>%
  summarize(across(.cols = everything(),
                  .fns = ~sum(is.na(.x)))) %>%
  t()

```


## EDA

**Data quality Report**

```{r}
skimr::skim(train)
```

*Potential Variables to retain in the analysis*

* start_speed (speed at 0 - 50ft)
* end_speed (speed at the front of home plate)
* pfx_x (a measure of break of the pitch in the x-direction)
* pfx_z (a measure of break of the pitch in the y-direction)
* px (a measure of x-value of position at the front of home plate)
* pz (a measure of the z-value of position at the front of home plate)
* x0 (x position at 50ft)
* z0 (z position at 50ft)
* vx0 (x velocity at 50ft)
* vz0 (z velocity at 50ft)

**How many pitch types are there?**

* Remove NA
* Remove KN, IN, PO, EP, and FO

```{r}
train %>%
  count(pitch_type, sort = TRUE)

```

*Pitch Types*

* FF = four-seam fastball
* SL = slider
* CH = change up
* FT = two-seam fastball
* SI = sinker
* CU = curveball
* FC = cutter
* KC = knuckle-curve
* FS = split-finger fastball

**Boxplots of our different variables by pitch type**

* Use the cleaned data, now

```{r}

pitch_boxplot_func <- function(x_var){
  train_cleaned %>%
    ggplot(aes(x = x_var, y = reorder(pitch_type, x_var))) +
    geom_boxplot()
}


pitch_boxplot_func(x_var = train_cleaned$start_speed)
pitch_boxplot_func(x_var = train_cleaned$end_speed)
pitch_boxplot_func(x_var = train_cleaned$pfx_x)
pitch_boxplot_func(x_var = train_cleaned$pfx_z)
pitch_boxplot_func(x_var = train_cleaned$px)
pitch_boxplot_func(x_var = train_cleaned$pz)
pitch_boxplot_func(x_var = train_cleaned$x0)
pitch_boxplot_func(x_var = train_cleaned$z0)
pitch_boxplot_func(x_var = train_cleaned$vx0)
pitch_boxplot_func(x_var = train_cleaned$vz0)

```


**NOTE:** `x0` and `z0` don't seem to have much variability between pitches, so they might not be that valuable.

**Correlation Between Variables**

```{r}

pitch_cor <- round(cor(train_cleaned[, -1]), 3)
pitch_cor

corrplot.mixed(pitch_cor, lower.col = "black", tl.cex = 0.7)

```


## Hierarchical Cluster Analysis

**Create a table of the median value for each pitch type**

```{r}

pitch_type_summary <- train_cleaned %>%
  group_by(pitch_type) %>%
  summarize(across(everything(),
            .fns = median))

```



**Create a vector of labels**

```{r}
pitches <- pitch_type_summary$pitch_type
```


**Create a table of cluster data**

```{r}

clust_df <- pitch_type_summary %>% select(-pitch_type) %>% as.data.frame()
rownames(clust_df) <- pitches

```


**Create Clusters**

```{r}
## scale the data and then use euclidean distance to identify "closeness"
pitch_clust <- dist(scale(clust_df))

## select a disimilarity metric
pitch_clust_final <- hclust(pitch_clust, method = "ward.D")

```


**Plot the dendogram**

```{r}
plot(pitch_clust_final, labels = pitches, main = "pitchf/x clusters")
```


Looks like about 4 clusters. Let's make those more distinct

```{r}
plot(pitch_clust_final, labels = pitches, main = "pitchf/x clusters")
rect.hclust(pitch_clust_final, k = 5, border = "red")
```


**Assign the clusters back to the data frame**

```{r}

clust_assignment <- cutree(pitch_clust_final, k = 4)

pitch_clust_df <- bind_cols(pitch_type = pitches, 
                            clust_df, 
                            cluster_num = clust_assignment)

## Which pitches in each cluster
pitch_clust_df %>%
  group_by(cluster_num) %>%
  count(pitch_type) %>%
  select(-n)

## summarize clusters
pitch_clust_df %>%
  group_by(cluster_num) %>%
  summarize(across(.cols = start_speed:vz0,
                  list(avg = mean, sd = sd))) %>%
  pivot_longer(-cluster_num) %>%
  pivot_wider(names_from = cluster_num,
              values_from = value)



```


**Plot the dendrogram in a more interesting way with {plotly}**

```{r}

dendro <- as.dendrogram(pitch_clust_final)

plot_dendro(dendro, heigh = 600,xmin = -10) %>%
  hide_legend() %>%
  highlight("plotly_selected", persistent = TRUE, dynamic = TRUE)


```

**Altneratively, conduct the clustering over all of the pitch in the data set**

```{r}

## pitch type labels
pitches <- train_cleaned$pitch_type

## data frame of cluster variables
clust_df <- train_cleaned %>% select(-pitch_type) %>% as.data.frame()

## scale the data and then use euclidean distance to identify "closeness"
pitch_clust <- dist(scale(clust_df))

## select a disimilarity metric
pitch_clust_final <- hclust(pitch_clust, method = "ward.D")

## get the cluster assignments
clust_assignment <- cutree(pitch_clust_final, k = 4)

## put the results in a data frame
pitch_clust_df <- bind_cols(pitch_type = pitches,
                            clust_df,
                            cluster_num = clust_assignment)

pitch_clust_df %>% dim()
pitch_clust_df %>% head()


## summarize clusters
pitch_clust_df %>%
  select(-pitch_type) %>%
  group_by(cluster_num) %>%
  summarize(across(.cols = start_speed:vz0,
                  list(avg = mean, sd = sd))) %>%
  pivot_longer(-cluster_num) %>%
  pivot_wider(names_from = cluster_num,
              values_from = value)

## Which pitches in each cluster
table("Pitch Type" = pitch_clust_df$pitch_type, "Cluster Number" = pitch_clust_df$cluster_num)

pitch_clust_df %>%
  count(cluster_num, pitch_type) %>%
  mutate(group = case_when(pitch_type %in% c("FF", "FT", "FC", "SI", "FS") ~ "Fastball Pitches",
                           TRUE ~ as.character("Off-Speed Pitches"))) %>%
  ggplot(aes(x = as.factor(cluster_num), y = n, fill = group)) +
  geom_col()


```






