---
title: "TidyX Episode 113 - R Classes - Making An S4 Object"
output: html_document
editor_options: 
  chunk_output_type: console
---

## R Objects

- Rich Object system (Base, S3, S4, R6, RC, ...R7?)
- Large % of R packages use S3
- Bioconductor embraced S4

See Hadley Wickham's thoughts in Advanced R -
https://adv-r.hadley.nz/oo.html

## Benefits of an S4 Object

- More strictly defined system
  - class defined name and architecture
- formalized "Slots" for storing data in the object
  - slot types will not change
- Uses "methods" still

"A (Not So) Short Introduction to S4" - Christophe Genolini https://cran.r-project.org/doc/contrib/Genolini-S4tutorialV0-5en.pdf


## Creating an S4

- S3 may be initialize with `structure()` function
- S4 requires object definition first with `setClass()`, then creation of a new object with `new()`

```{r}

## define the class
setClass(
  Class = "my_s4",
  ## what sort of data can be entered
  slots = c(x = "numeric"),
  ## default values
  prototype = list(x = numeric())
  )

my_s4_obj <- new("my_s4", x = 24) 

my_s4_obj

#invalid entry into x
new("my_s4", x = "asdofnasodinf") 

```

## Defining your Object

- use "class" tag to add a _new_ class
- set various internal attributes using other tags
- Add data/non meta information via .Data

```{r}

setClass(
  Class = "custom_message",
  ## what sort of data can be entered
  slots = c(message = "character",
            authors = "character",
            creation_time = "POSIXct"), 
  ## default values
  prototype = list(
    message = character(),
    authors = character(),
    creation_time = Sys.time()
  )
)

my_message <- new(
  "custom_message",
  message = "TidyX is Awesome!",
  authors = c("Ellis", "Patrick")
  )

my_message

```

## Developing Methods

- print, format, etc are all "generics", applying a method developed for the specific class
- ie. predict across all the model types!
- s3 uses dot notation to create new methods: print.custom_message
- s4 methods must be set using `setMethod()`, and uses the "show" method

```{r}

setMethod("show",signature = signature(object = "custom_message"),
  function(object){
    print_message <- c(
      "< Start Message >",
      paste0("  Contents: ",object@message),
      paste0("  Author: ", paste0(object@authors, collapse = ", ")),
      paste0("  Creation Time: ", format(object@creation_time)),
      "< End Message >"
    )
    
    cat(
      print_message,
      sep = "\n"
    )
  })

my_message

```

## Develop Class Creator

- Avoid calling `new()` directly
- Abstract some of the concepts needed

```{r}

new_message <- function(message, by){
  new(
    "custom_message",
    message = message,
    authors = by
  )
}

m1 <- new_message(
  "You should Like and Subscribe to TidyX",
  by = "Ellis"
)

m1

new_message(
  "Consider becoming a Patron at Patreon - patreon.com/Tidy_Explained",
  by = c("Patrick", "Ellis")
)

```


