---
title: "TidyX Episode 130 - MLB Hitting Production"
author: "TidyX"
output: 
  html_document:
    toc: TRUE
    number_sections: FALSE
    toc_float: TRUE
    toc_collapsed: FALSE
    theme: lumen
editor_options: 
  chunk_output_type: console
---

# Report Details {.tabset}

**This report provides a view into the hitting performance of MLB players drafted 2010 and on who have played in at least 8 seasons.**

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)

suppressWarnings({
  suppressPackageStartupMessages({
    library(crosstalk)
    library(plotly)
    library(tidyverse)
    library(Lahman)
  })
})

```

```{r data-prep, include = FALSE}

### Data
dat <- Batting %>%
  #Person information
  left_join(People) %>% 
  # Get the position played by the person that year
  left_join(Fielding %>% select(playerID, yearID, POS)) %>% 
  mutate(
    first_season = substr(debut, start = 1, stop = 4), # Had a full date 
    name = paste(nameLast, nameFirst, sep = ", ")
  ) %>%
  filter(first_season >= 2010) %>%
  mutate(
    obp = (H + BB + HBP) / (AB + BB + HBP + SF),
    batting_avg = H / AB
  ) %>%
  select(playerID, name, yearID, POS, AB, batting_avg, obp) %>%
  filter(AB >= 100) %>%
  group_by(playerID) %>%
  mutate(experience = row_number()) %>%
  arrange(playerID) %>%
  ## Reduce number of players listed to only those that played 8 seasons (arbitrary)
  group_by(playerID) %>%
  filter((max(yearID) - min(yearID) >= 7)) ## Min 8 years (rookie year is year 0)

```

## Player Production

### Rookie Season Production

-   Select the rookie season of interest
-   Place your cursor over a dot to see which player it is

```{r}
#### Rookie season performance plot
## set highlighting capabilities for rookie season stats
rookie_key <- highlight_key(
  x = dat %>% filter(experience == 1),
  ~yearID
  )

rookie_plt <- bscols(
  widths = c(2,10),
  filter_checkbox(
    id = "yearID", 
    label = "Season",
    sharedData = rookie_key,
    group =  ~yearID
    ),
  
  plot_ly(
    data = rookie_key, 
    x = ~batting_avg, 
    y = ~obp, 
    showlegend = FALSE
    ) %>%
    ## add points and the text argument adds tooltip information
    add_markers(
      text = ~name,
      marker = list(color = '#56B4E9', size = 15)
    ) %>% 
    layout(yaxis = list(title = "OBP")) %>%
    layout(xaxis = list(title = "Batting Avg")) %>%
    layout(title = "Rookie Season Production")
)

rookie_plt

```

### Player Time Series

-   Select a player of interest to observe their trend relative to the population
-   Add additional players to compare their performance and change color so that the individuals stand out

```{r}
#### Player time series plot
# player highlight key
player_key <- highlight_key(dat, ~name)

# player time series plot using the highlight key
player_ts <- plot_ly(
    data = player_key, 
    color = I("black"),
    line = list(width = 5)
  ) %>%
  group_by(name) %>%
  add_lines(
    x = ~yearID,
    y = ~batting_avg
  ) %>% 
  layout(yaxis = list(title = "Batting Average"), autosize = FALSE) %>%
  layout(xaxis = list(title = "Season"), autosize = FALSE) %>%
  layout(title = "Career Batting Average Trend")   

# player plot with interactive capability
player_plt <- highlight(
  player_ts,
  selectize = TRUE,
  dynamic = TRUE,
  persistent = TRUE
)

player_plt
```

## Distribution of performance by position

-   Select the position of interest to see on base percentage
-   Change colors when adding new positions to allow for comparison

```{r}
#### Position histogram
# position highlight key
pos_key <- highlight_key(dat, ~POS)

# BA histogram
ba_hist <- plot_ly(
  data = pos_key, 
  color = I("light grey")
  ) %>%
  group_by(POS) %>%
  add_histogram(x = ~batting_avg) %>%
  layout(
    xaxis = list(title = "Batting Avg"), 
    autosize = FALSE
    )

# OBP histogram
obp_hist <- plot_ly(
    data = pos_key, 
    color = I("light grey")
  ) %>%
  group_by(POS) %>%
  add_histogram(x = ~obp) %>%
  layout(
    xaxis = list(title = "OBP"), 
    autosize = FALSE
    )

subplot(
  ba_hist, 
  obp_hist, 
  nrows = 1, 
  shareY = TRUE,
  titleX = TRUE
  ) %>%
  layout(
    barmode = "overlay",
    showlegend = FALSE
  ) %>%
  highlight(
    dynamic = TRUE,
    selectize = TRUE,
    persistent = TRUE,
    selected = attrs_selected(opacity = 0.5)
  )

```
