---
title: "TidyX Episode 114 - camcorder"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

{camcorder} has been a fun project for the past \~ 2 years

<https://github.com/thebioengineer/camcorder>

Developed to help Cédric Scherer and Georgios Karamanis with their cool gifs that are recordings of their plot making process

Recently highlighted by Rami Krispin (<https://www.linkedin.com/feed/update/urn:li:activity:6953022051725041664/>)

## Thoughts behind package

-   Plot making is an iterative process
-   Tools in R to make plots are super powerful
-   *but* the rstudio "viewer" and "plots" panes are not ideal
-   Manually saving the image to open and view. or change sizes is not simple

```{r}

library(ggplot2)

ggplot(cars) + 
  geom_point(aes(x = speed, y = dist))

## To see actual output, need to save and view file
ggsave(file.path(tempdir(),"examplefile.pdf"), device = cairo_pdf)

## only works on windows computers(!!)
shell.exec(file.path(tempdir(),"examplefile.pdf"))

unlink(file.path(tempdir(),"examplefile.pdf"))

```

## Camcorder

```{r}

library(camcorder)

## to initalize recording

gg_record(device = cairo_pdf)

ggplot(cars) + 
  geom_point(aes(x = speed, y = dist))


ggplot(cars) + 
  geom_point(aes(x = speed, y = dist)) + 
  labs(
    title = "Car Speed vs Stopping Distance"
  )
  
ggplot(cars) + 
  geom_point(aes(x = speed, y = dist)) +
  geom_vline(xintercept = 20) +
  labs(
    title = "Car Speed vs Stopping Distance"
  )

ggplot(cars) + 
  geom_point(aes(x = speed, y = dist)) +
  geom_smooth(aes(x = speed, y = dist)) +
  geom_vline(xintercept = 20) +
  labs(
    title = "Car Speed vs Stopping Distance"
  )

ggplot(cars) + 
  geom_point(aes(x = speed, y = dist)) +
  geom_smooth(aes(x = speed, y = dist)) +
  geom_vline(xintercept = 20) +
  labs(
    title = "Car Speed vs Stopping Distance",
    x = "Speed (MPH)",
    y = "Stopping Distance (ft)"
  ) 


ggplot(cars) + 
  geom_point(aes(x = speed, y = dist)) +
  geom_smooth(aes(x = speed, y = dist)) +
  geom_vline(xintercept = 20) +
  labs(
    title = "Car Speed vs Stopping Distance",
    x = "Speed (MPH)",
    y = "Stopping Distance (ft)"
  ) +
  theme_classic(
    base_family = "Times New Roman"
  )

gg_stop_recording()



```

```{r}

gg_playback(first_image_duration = 3) ## sets first image to be displayed for 3 frames

```
